---
title: "Applying Hessilt et al. (2022) Methods"
author: "Charlotte Uden"
date: "2025-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(patchwork)
library(rnaturalearth)
library(rnaturalearthdata)
library(glmnet)
library(caret) 
library(pROC)
library(sf)
library(gridExtra)
library(dplyr)

#for bayes models:
library(rstan)
library(pROC)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]
```

############################################
############################################
# GLMNET Penalized logistic regression 
############################################
############################################

#----------------------------
## Data for 2002 - 2011
#----------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2002_2011_10x.csv")
df <- read.csv(drivers_file_path)[,2:13]
df <- na.omit(df) 

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```

Function to apply the penalized logistic ridge regression to lightning caused fire data

inputs:
1. dataset
2. vector of predictor variables you would like to apply
3. response variable
4. alpha (ridge or lass0)
5. fraction of data to put in the train set
6. seed to be set to (keep this the same for all models)
7. number of folds for cross validation

funcitons:
computes training means/SDs
stores them in the output list
applies them to training and test sets
fits glmnet with standardize = FALSE

returns:
1. coefficents
2. predictions
3. perfomance metrics
4. standardization values (means and sd for each predictor) for future model application
4. a helper function to standardize future data identically


```{r}
#library(glmnet)
#library(pROC)

library(glmnet)
library(pROC)

fit_penalized_logit <- function(data,
                                predictors,
                                response = "caused_fire",
                                alpha = 0,          # 0 = ridge, 1 = lasso
                                train_frac = 0.8,
                                seed = 123,
                                nfolds = 5,
                                boot = TRUE,
                                n_boot = 1000) {
  
  # ---- helper: PR-AUC (average precision style) ---------------------------
  pr_auc_ap <- function(y_true, p_hat) {
    o <- order(p_hat, decreasing = TRUE)
    y <- y_true[o]
    tp <- cumsum(y == 1)
    fp <- cumsum(y == 0)
    total_pos <- sum(y == 1)
    if (total_pos == 0) return(NA_real_)
    recall <- tp / total_pos
    precision <- tp / (tp + fp)
    recall_prev <- c(0, recall[-length(recall)])
    sum((recall - recall_prev) * precision)
  }
  
  # ---- helper: calibration intercept & slope ------------------------------
  calibration_stats <- function(y_true, p_hat) {
    eps <- 1e-6
    p_clipped <- pmin(pmax(p_hat, eps), 1 - eps)
    logit_p <- qlogis(p_clipped)
    cal_fit <- glm(y_true ~ logit_p, family = binomial())
    coefs <- coef(cal_fit)
    list(
      intercept = unname(coefs[1]),
      slope     = unname(coefs["logit_p"])
    )
  }
  
  # 1. Train/test split -------------------------------------------------------
  set.seed(seed)
  n <- nrow(data)
  train_size <- floor(train_frac * n)
  train_index <- sample(seq_len(n), size = train_size)
  
  train <- data[train_index, , drop = FALSE]
  test  <- data[-train_index, , drop = FALSE]
  
  # 2. Compute scaling values (mean/sd) from training data -------------------
  scale_values <- train[, predictors, drop = FALSE] |>
    lapply(function(x) c(mean = mean(x), sd = sd(x))) |>
    do.call(what = rbind)
  
  # Standardization helper
  standardize_df <- function(df) {
    out <- df
    for (p in predictors) {
      m <- scale_values[p, "mean"]
      s <- scale_values[p, "sd"]
      out[[p]] <- (out[[p]] - m) / s
    }
    out
  }
  
  train_std <- standardize_df(train)
  test_std  <- standardize_df(test)
  
  # 3. Build model matrices ---------------------------------------------------
  formula_str <- paste(response, "~", paste(predictors, collapse = " + "))
  formula_obj <- as.formula(formula_str)
  
  x_train <- model.matrix(formula_obj, data = train_std)[, -1, drop = FALSE]
  y_train <- train_std[[response]]
  
  x_test  <- model.matrix(formula_obj, data = test_std)[, -1, drop = FALSE]
  y_test  <- test_std[[response]]
  
  # 4. Fit penalized logistic regression -------------------------------------
  cv_fit <- cv.glmnet(
    x = x_train,
    y = y_train,
    family = "binomial",
    alpha = alpha,
    nfolds = nfolds,
    type.measure = "deviance",
    standardize = FALSE   # we already did it
  )
  
  best_lambda <- cv_fit$lambda.1se
  
  # 5. Extract coefficients ---------------------------------------------------
  coef_mat <- coef(cv_fit, s = best_lambda)
  coef_df <- data.frame(
    term = rownames(coef_mat),
    estimate = as.numeric(coef_mat),
    row.names = NULL
  )
  
  # 6. Predictions ------------------------------------------------------------
  prob_test <- as.numeric(
    predict(cv_fit, newx = x_test, s = best_lambda, type = "response")
  )
  class_test <- ifelse(prob_test >= 0.5, 1L, 0L)
  
  # 7. Point-estimate performance metrics ------------------------------------
  roc_obj <- roc(response = y_test, predictor = prob_test)
  auc_val <- as.numeric(auc(roc_obj))
  
  accuracy <- mean(class_test == y_test)
  brier <- mean((prob_test - y_test)^2)
  pr_auc_val <- pr_auc_ap(y_test, prob_test)
  cal <- calibration_stats(y_test, prob_test)
  
  # 8. Bootstrap CIs (optional) ----------------------------------------------
  AUC_CI <- PR_AUC_CI <- ACC_CI <- Brier_CI <- 
    calib_int_CI <- calib_slope_CI <- c(NA_real_, NA_real_)
  
  if (boot) {
    n_test <- length(y_test)
    AUC_boot   <- numeric(n_boot)
    PR_boot    <- numeric(n_boot)
    ACC_boot   <- numeric(n_boot)
    Brier_boot <- numeric(n_boot)
    INT_boot   <- numeric(n_boot)
    SLOPE_boot <- numeric(n_boot)
    
    for (b in seq_len(n_boot)) {
      idx <- sample.int(n_test, n_test, replace = TRUE)
      y_b <- y_test[idx]
      p_b <- prob_test[idx]
      c_b <- ifelse(p_b >= 0.5, 1L, 0L)
      
      # AUC
      if (length(unique(y_b)) == 2) {
        AUC_boot[b] <- tryCatch(
          as.numeric(auc(roc(response = y_b, predictor = p_b))),
          error = function(e) NA_real_
        )
      } else {
        AUC_boot[b] <- NA_real_
      }
      
      # PR-AUC
      PR_boot[b] <- pr_auc_ap(y_b, p_b)
      
      # Accuracy
      ACC_boot[b] <- mean(c_b == y_b)
      
      # Brier
      Brier_boot[b] <- mean((p_b - y_b)^2)
      
      # Calibration
      cal_b <- tryCatch(
        calibration_stats(y_b, p_b),
        error = function(e) list(intercept = NA_real_, slope = NA_real_)
      )
      INT_boot[b]   <- cal_b$intercept
      SLOPE_boot[b] <- cal_b$slope
    }
    
    qfun <- function(x) quantile(x[is.finite(x) & !is.na(x)], probs = c(0.025, 0.975))
    
    if (any(is.finite(AUC_boot)))       AUC_CI       <- qfun(AUC_boot)
    if (any(is.finite(PR_boot)))        PR_AUC_CI    <- qfun(PR_boot)
    if (any(is.finite(ACC_boot)))       ACC_CI       <- qfun(ACC_boot)
    if (any(is.finite(Brier_boot)))     Brier_CI     <- qfun(Brier_boot)
    if (any(is.finite(INT_boot)))       calib_int_CI <- qfun(INT_boot)
    if (any(is.finite(SLOPE_boot)))     calib_slope_CI <- qfun(SLOPE_boot)
  }
  
  # 9. Return everything ------------------------------------------------------
  list(
    call = match.call(),
    formula = formula_obj,
    predictors = predictors,
    alpha = alpha,
    lambda = best_lambda,
    coefficients = coef_df,
    scale_values = scale_values,
    cv_fit = cv_fit,
    predictions = data.frame(
      observed = y_test,
      predicted_prob = prob_test,
      predicted_class = class_test
    ),
    performance = list(
      AUC = auc_val,
      AUC_CI = AUC_CI,
      PR_AUC = pr_auc_val,
      PR_AUC_CI = PR_AUC_CI,
      accuracy = accuracy,
      accuracy_CI = ACC_CI,
      Brier = brier,
      Brier_CI = Brier_CI,
      calib_intercept = cal$intercept,
      calib_intercept_CI = calib_int_CI,
      calib_slope = cal$slope,
      calib_slope_CI = calib_slope_CI
    ),
    standardize_newdata = function(newdata) {
      out <- newdata
      for (p in predictors) {
        m <- scale_values[p, "mean"]
        s <- scale_values[p, "sd"]
        out[[p]] <- (out[[p]] - m) / s
      }
      out
    }
  )
}


```

M1: climate only

```{r}
predictors_m1 <- c("temperature", "precipitation", "wind",
                   "relative_humidity")

m1 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m1,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M2: fuel moisture only

```{r}
predictors_m2 <- c("FFMC", "DMC", "DC")

m2 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m2,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M3: Vegetation type/cover only

```{r}
predictors_m3 <- c("C", "B", "U")

m3 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m3,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m3$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M4: Climate and fuel moisture

```{r}
predictors_m4 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC")

m4 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m4,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m4$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```


M5: Climate and vegetation cover

```{r}
predictors_m5 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "C", "B", "U")

m5 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m5,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m5$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M6: fuel moisture and vegetation cover

```{r}
predictors_m6 <- c("FFMC", "DMC", "DC", 
                   "C", "B", "U")

m6 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m6,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m6$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M7: Climate, fuel moisture, and vegetation type/cover

```{r}
predictors_m7 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")

m7 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m7,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m7$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

Compare Performance

```{r}
model <- c("m1", "m2", "m3", "m4", "m5", "m6", "m7")
training_years <- rep("2001-2012", 7)
approach <- rep("glmnet", 7)
predictors <- c("climate", "fuel moisture", "vegetation", "climate & fuel moisture", "climate & vegetation", "fuel moisture & vegetation", "climate, fuel moisture & vegetation")

AUC <- c(m1$performance$AUC, m2$performance$AUC, m3$performance$AUC, m4$performance$AUC, m5$performance$AUC, m6$performance$AUC, m7$performance$AUC)

PR_AUC <- c(m1$performance$PR_AUC, m2$performance$PR_AUC, m3$performance$PR_AUC, m4$performance$PR_AUC, m5$performance$PR_AUC, m6$performance$PR_AUC, m7$performance$PR_AUC)

accuracy <- c(m1$performance$accuracy, m2$performance$accuracy, m3$performance$accuracy, m4$performance$accuracy, m5$performance$accuracy, m6$performance$accuracy, m7$performance$accuracy)

Brier <- c(m1$performance$Brier, m2$performance$Brier, m3$performance$Brier, m4$performance$Brier, m5$performance$Brier, m6$performance$Brier, m7$performance$Brier)

calib_intercept <- c(m1$performance$calib_intercept, m2$performance$calib_intercept, m3$performance$calib_intercept, m4$performance$calib_intercept, m5$performance$calib_intercept, m6$performance$calib_intercept, m7$performance$calib_intercept)

calib_slope <- c(m1$performance$calib_slope, m2$performance$calib_slope, m3$performance$calib_slope, m4$performance$calib_slope, m5$performance$calib_slope, m6$performance$calib_slope, m7$performance$calib_slope)

#confidence intervals
AUC_CI_low <- c(m1$performance$AUC_CI[1], m2$performance$AUC_CI[1], m3$performance$AUC_CI[1], m4$performance$AUC_CI[1], m5$performance$AUC_CI[1], m6$performance$AUC_CI[1], m7$performance$AUC_CI[1])
AUC_CI_high <- c(m1$performance$AUC_CI[2], m2$performance$AUC_CI[2], m3$performance$AUC_CI[2], m4$performance$AUC_CI[2], m5$performance$AUC_CI[2], m6$performance$AUC_CI[2], m7$performance$AUC_CI[2])

PR_AUC_CI_low <- c(m1$performance$PR_AUC_CI[1], m2$performance$PR_AUC_CI[1], m3$performance$PR_AUC_CI[1], m4$performance$PR_AUC_CI[1], m5$performance$PR_AUC_CI[1], m6$performance$PR_AUC_CI[1], m7$performance$PR_AUC_CI[1])
PR_AUC_CI_high <- c(m1$performance$PR_AUC_CI[2], m2$performance$PR_AUC_CI[2], m3$performance$PR_AUC_CI[2], m4$performance$PR_AUC_CI[2], m5$performance$PR_AUC_CI[2], m6$performance$PR_AUC_CI[2], m7$performance$PR_AUC_CI[2])

Brier_CI_low <- c(m1$performance$Brier_CI[1], m2$performance$Brier_CI[1], m3$performance$Brier_CI[1], m4$performance$Brier_CI[1], m5$performance$Brier_CI[1], m6$performance$Brier_CI[1], m7$performance$Brier_CI[1])
Brier_CI_high <- c(m1$performance$Brier_CI[2], m2$performance$Brier_CI[2], m3$performance$Brier_CI[2], m4$performance$Brier_CI[2], m5$performance$Brier_CI[2], m6$performance$Brier_CI[2], m7$performance$Brier_CI[2])

calib_intercept_CI_low <- c(m1$performance$calib_intercept_CI[1], m2$performance$calib_intercept_CI[1], m3$performance$calib_intercept_CI[1], m4$performance$calib_intercept_CI[1], m5$performance$calib_intercept_CI[1], m6$performance$calib_intercept_CI[1], m7$performance$calib_intercept_CI[1])
calib_intercept_CI_high <- c(m1$performance$calib_intercept_CI[2], m2$performance$calib_intercept_CI[2], m3$performance$calib_intercept_CI[2], m4$performance$calib_intercept_CI[2], m5$performance$calib_intercept_CI[2], m6$performance$calib_intercept_CI[2], m7$performance$calib_intercept_CI[2])

calib_slope_CI_low <- c(m1$performance$calib_slope_CI[1], m2$performance$calib_slope_CI[1], m3$performance$calib_slope_CI[1], m4$performance$calib_slope_CI[1], m5$performance$calib_slope_CI[1], m6$performance$calib_slope_CI[1], m7$performance$calib_slope_CI[1])
calib_slope_CI_high <- c(m1$performance$calib_slope_CI[2], m2$performance$calib_slope_CI[2], m3$performance$calib_slope_CI[2], m4$performance$calib_slope_CI[2], m5$performance$calib_slope_CI[2], m6$performance$calib_slope_CI[2], m7$performance$calib_slope_CI[2])

df_performance_1 <- as.data.frame(cbind(model, training_years, approach, predictors, AUC, PR_AUC, accuracy, Brier, calib_intercept, calib_slope, AUC_CI_low, AUC_CI_high, PR_AUC_CI_low, PR_AUC_CI_high, Brier_CI_low, Brier_CI_high, calib_intercept_CI_low, calib_intercept_CI_high, calib_slope_CI_low, calib_slope_CI_high))

p1 <- ggplot(data=df_performance_1, aes(y=AUC, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p2 <- ggplot(data=df_performance_1, aes(y=accuracy, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p3 <- ggplot(data=df_performance_1, aes(y=Brier, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p1 + p2 + p3
```

#----------------------------
## Data for 2012 - 2018
#----------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2012_2018_10x.csv")
df <- read.csv(drivers_file_path)

df <- na.omit(df) 

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```

M1: climate only

```{r}
predictors_m1 <- c("temperature", "precipitation", "wind",
                   "relative_humidity")

m1 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m1,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M2: fuel moisture only

```{r}
predictors_m2 <- c("FFMC", "DMC", "DC")

m2 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m2,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M3: Vegetation type/cover only

```{r}
predictors_m3 <- c("C", "B", "U")

m3 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m3,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m3$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```


M4: Climate and fuel moisture

```{r}
predictors_m4 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC")

m4 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m4,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m4$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```


M5: Climate and vegetation cover

```{r}
predictors_m5 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "C", "B", "U")

m5 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m5,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m5$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M6: fuel moisture and vegetation cover

```{r}
predictors_m6 <- c("FFMC", "DMC", "DC", 
                   "C", "B", "U")

m6 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m6,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m6$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

M7: Climate, fuel moisture, and vegetation type/cover

```{r}
predictors_m7 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")

m7 <- fit_penalized_logit(
  data       = df,
  predictors = predictors_m7,
  response   = "caused_fire",
  alpha      = 0  # ridge, like Hessilt et al.
)

ggplot() + geom_point(data=m7$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("Observed vs. Predicted") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

```

Compare models

```{r}
model <- c("m1", "m2", "m3", "m4", "m5", "m6", "m7")
training_years <- rep("2012-2018", 7)
approach <- rep("glmnet", 7)
predictors <- c("climate", "fuel moisture", "vegetation", "climate & fuel moisture", "climate & vegetation", "fuel moisture & vegetation", "climate, fuel moisture & vegetation")

AUC <- c(m1$performance$AUC, m2$performance$AUC, m3$performance$AUC, m4$performance$AUC, m5$performance$AUC, m6$performance$AUC, m7$performance$AUC)

PR_AUC <- c(m1$performance$PR_AUC, m2$performance$PR_AUC, m3$performance$PR_AUC, m4$performance$PR_AUC, m5$performance$PR_AUC, m6$performance$PR_AUC, m7$performance$PR_AUC)

accuracy <- c(m1$performance$accuracy, m2$performance$accuracy, m3$performance$accuracy, m4$performance$accuracy, m5$performance$accuracy, m6$performance$accuracy, m7$performance$accuracy)

Brier <- c(m1$performance$Brier, m2$performance$Brier, m3$performance$Brier, m4$performance$Brier, m5$performance$Brier, m6$performance$Brier, m7$performance$Brier)

calib_intercept <- c(m1$performance$calib_intercept, m2$performance$calib_intercept, m3$performance$calib_intercept, m4$performance$calib_intercept, m5$performance$calib_intercept, m6$performance$calib_intercept, m7$performance$calib_intercept)

calib_slope <- c(m1$performance$calib_slope, m2$performance$calib_slope, m3$performance$calib_slope, m4$performance$calib_slope, m5$performance$calib_slope, m6$performance$calib_slope, m7$performance$calib_slope)

#confidence intervals
AUC_CI_low <- c(m1$performance$AUC_CI[1], m2$performance$AUC_CI[1], m3$performance$AUC_CI[1], m4$performance$AUC_CI[1], m5$performance$AUC_CI[1], m6$performance$AUC_CI[1], m7$performance$AUC_CI[1])
AUC_CI_high <- c(m1$performance$AUC_CI[2], m2$performance$AUC_CI[2], m3$performance$AUC_CI[2], m4$performance$AUC_CI[2], m5$performance$AUC_CI[2], m6$performance$AUC_CI[2], m7$performance$AUC_CI[2])

PR_AUC_CI_low <- c(m1$performance$PR_AUC_CI[1], m2$performance$PR_AUC_CI[1], m3$performance$PR_AUC_CI[1], m4$performance$PR_AUC_CI[1], m5$performance$PR_AUC_CI[1], m6$performance$PR_AUC_CI[1], m7$performance$PR_AUC_CI[1])
PR_AUC_CI_high <- c(m1$performance$PR_AUC_CI[2], m2$performance$PR_AUC_CI[2], m3$performance$PR_AUC_CI[2], m4$performance$PR_AUC_CI[2], m5$performance$PR_AUC_CI[2], m6$performance$PR_AUC_CI[2], m7$performance$PR_AUC_CI[2])

Brier_CI_low <- c(m1$performance$Brier_CI[1], m2$performance$Brier_CI[1], m3$performance$Brier_CI[1], m4$performance$Brier_CI[1], m5$performance$Brier_CI[1], m6$performance$Brier_CI[1], m7$performance$Brier_CI[1])
Brier_CI_high <- c(m1$performance$Brier_CI[2], m2$performance$Brier_CI[2], m3$performance$Brier_CI[2], m4$performance$Brier_CI[2], m5$performance$Brier_CI[2], m6$performance$Brier_CI[2], m7$performance$Brier_CI[2])

calib_intercept_CI_low <- c(m1$performance$calib_intercept_CI[1], m2$performance$calib_intercept_CI[1], m3$performance$calib_intercept_CI[1], m4$performance$calib_intercept_CI[1], m5$performance$calib_intercept_CI[1], m6$performance$calib_intercept_CI[1], m7$performance$calib_intercept_CI[1])
calib_intercept_CI_high <- c(m1$performance$calib_intercept_CI[2], m2$performance$calib_intercept_CI[2], m3$performance$calib_intercept_CI[2], m4$performance$calib_intercept_CI[2], m5$performance$calib_intercept_CI[2], m6$performance$calib_intercept_CI[2], m7$performance$calib_intercept_CI[2])

calib_slope_CI_low <- c(m1$performance$calib_slope_CI[1], m2$performance$calib_slope_CI[1], m3$performance$calib_slope_CI[1], m4$performance$calib_slope_CI[1], m5$performance$calib_slope_CI[1], m6$performance$calib_slope_CI[1], m7$performance$calib_slope_CI[1])
calib_slope_CI_high <- c(m1$performance$calib_slope_CI[2], m2$performance$calib_slope_CI[2], m3$performance$calib_slope_CI[2], m4$performance$calib_slope_CI[2], m5$performance$calib_slope_CI[2], m6$performance$calib_slope_CI[2], m7$performance$calib_slope_CI[2])

df_performance_2 <- as.data.frame(cbind(model, training_years, approach, predictors, AUC, PR_AUC, accuracy, Brier, calib_intercept, calib_slope, AUC_CI_low, AUC_CI_high, PR_AUC_CI_low, PR_AUC_CI_high, Brier_CI_low, Brier_CI_high, calib_intercept_CI_low, calib_intercept_CI_high, calib_slope_CI_low, calib_slope_CI_high))

p1 <- ggplot(data=df_performance_2, aes(y=AUC, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p2 <- ggplot(data=df_performance_2, aes(y=accuracy, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Accuracy")

p3 <- ggplot(data=df_performance_2, aes(y=Brier, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Brier")

p1 + p2 + p3
```

############################################
############################################
# Bayesian Models
############################################
############################################

#----------------------------
## Data for 2002 - 2011
#----------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2002_2011_10x.csv")
df <- read.csv(drivers_file_path)

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```

```{r}

fit_bayes_logit_ridge <- function(data,
                                  predictors,
                                  response = "caused_fire",
                                  train_frac = 0.8,
                                  seed = 123,
                                  iter = 2000,
                                  chains = 4,
                                  sigma_beta = 1,
                                  sigma_alpha = 5,
                                  compiled_model = NULL,
                                  boot = TRUE,
                                  n_boot = 1000) {
  #-------------------------------
  # 0. Stan model code (inline)
  #-------------------------------
  stan_code <- "
  data {
    int<lower=1> N;
    int<lower=1> K;
    matrix[N, K] X;
    int<lower=0, upper=1> y[N];
    real<lower=0> sigma_beta;
    real<lower=0> sigma_alpha;
  }
  parameters {
    real alpha;
    vector[K] beta;
  }
  model {
    alpha ~ normal(0, sigma_alpha);
    beta  ~ normal(0, sigma_beta);
    y ~ bernoulli_logit(alpha + X * beta);
  }
  generated quantities {
    vector[N] log_lik;
    for (n in 1:N) {
      log_lik[n] = bernoulli_logit_lpmf(y[n] | alpha + X[n] * beta);
    }
  }"
  
  # ---- helper: PR-AUC (average precision style) ---------------------------
  pr_auc_ap <- function(y_true, p_hat) {
    o <- order(p_hat, decreasing = TRUE)
    y <- y_true[o]
    tp <- cumsum(y == 1)
    fp <- cumsum(y == 0)
    total_pos <- sum(y == 1)
    if (total_pos == 0) return(NA_real_)
    recall <- tp / total_pos
    precision <- tp / (tp + fp)
    recall_prev <- c(0, recall[-length(recall)])
    sum((recall - recall_prev) * precision)
  }
  
  # ---- helper: calibration intercept & slope ------------------------------
  calibration_stats <- function(y_true, p_hat) {
    eps <- 1e-6
    p_clipped <- pmin(pmax(p_hat, eps), 1 - eps)
    logit_p <- qlogis(p_clipped)
    cal_fit <- glm(y_true ~ logit_p, family = binomial())
    coefs <- coef(cal_fit)
    list(
      intercept = unname(coefs[1]),
      slope     = unname(coefs["logit_p"])
    )
  }
  
  #-------------------------------
  # 1. Train/test split
  #-------------------------------
  set.seed(seed)
  n <- nrow(data)
  train_size <- floor(train_frac * n)
  train_index <- sample(seq_len(n), size = train_size)
  
  train <- data[train_index, , drop = FALSE]
  test  <- data[-train_index, , drop = FALSE]
  
  #-------------------------------
  # 2. Standardization (training-based)
  #-------------------------------
  scale_list <- lapply(train[, predictors, drop = FALSE],
                       function(x) c(mean = mean(x), sd = sd(x)))
  scale_values <- do.call(rbind, scale_list)
  
  standardize_df <- function(df) {
    out <- df
    for (p in predictors) {
      m <- scale_values[p, "mean"]
      s <- scale_values[p, "sd"]
      out[[p]] <- (out[[p]] - m) / s
    }
    out
  }
  
  train_std <- standardize_df(train)
  test_std  <- standardize_df(test)
  
  #-------------------------------
  # 3. Model matrices
  #-------------------------------
  formula_str <- paste(response, "~", paste(predictors, collapse = " + "))
  formula_obj <- as.formula(formula_str)
  
  X_train <- model.matrix(formula_obj, data = train_std)[, -1, drop = FALSE]
  y_train <- train_std[[response]]
  
  X_test  <- model.matrix(formula_obj, data = test_std)[, -1, drop = FALSE]
  y_test  <- test_std[[response]]
  
  N <- nrow(X_train)
  K <- ncol(X_train)
  
  stan_data <- list(
    N = N,
    K = K,
    X = X_train,
    y = y_train,
    sigma_beta = sigma_beta,
    sigma_alpha = sigma_alpha
  )
  
  #-------------------------------
  # 4. Compile Stan model (once per session)
  #-------------------------------
  if (is.null(compiled_model)) {
    sm <- rstan::stan_model(model_code = stan_code)
  } else {
    sm <- compiled_model
  }
  
  #-------------------------------
  # 5. Sample from posterior
  #-------------------------------
  fit <- rstan::sampling(
    sm,
    data = stan_data,
    iter = iter,
    chains = chains,
    seed = seed,
    refresh = 0
  )
  
  #-------------------------------
  # 6. Posterior summaries (alpha, beta)
  #-------------------------------
  summ <- rstan::summary(fit, pars = c("alpha", "beta"))$summary
  alpha_mean <- summ["alpha", "mean"]
  beta_means <- summ[grep("^beta\\[", rownames(summ)), "mean"]
  names(beta_means) <- colnames(X_train)
  
  #-------------------------------
  # 7. Test-set predictions (using posterior means)
  #-------------------------------
  linpred_test <- as.vector(alpha_mean + X_test %*% beta_means)
  prob_test <- plogis(linpred_test)
  class_test <- ifelse(prob_test >= 0.5, 1L, 0L)
  
  #-------------------------------
  # 8. Point-estimate performance metrics
  #-------------------------------
  roc_obj <- roc(response = y_test, predictor = prob_test)
  auc_val <- as.numeric(auc(roc_obj))
  
  accuracy <- mean(class_test == y_test)
  brier <- mean((prob_test - y_test)^2)
  pr_auc_val <- pr_auc_ap(y_test, prob_test)
  cal <- calibration_stats(y_test, prob_test)
  
  #-------------------------------
  # 9. Bootstrap CIs (optional)
  #-------------------------------
  AUC_CI <- PR_AUC_CI <- ACC_CI <- Brier_CI <- 
    calib_int_CI <- calib_slope_CI <- c(NA_real_, NA_real_)
  
  if (boot) {
    n_test <- length(y_test)
    AUC_boot   <- numeric(n_boot)
    PR_boot    <- numeric(n_boot)
    ACC_boot   <- numeric(n_boot)
    Brier_boot <- numeric(n_boot)
    INT_boot   <- numeric(n_boot)
    SLOPE_boot <- numeric(n_boot)
    
    for (b in seq_len(n_boot)) {
      idx <- sample.int(n_test, n_test, replace = TRUE)
      y_b <- y_test[idx]
      p_b <- prob_test[idx]
      c_b <- ifelse(p_b >= 0.5, 1L, 0L)
      
      # AUC
      if (length(unique(y_b)) == 2) {
        AUC_boot[b] <- tryCatch(
          as.numeric(auc(roc(response = y_b, predictor = p_b))),
          error = function(e) NA_real_
        )
      } else {
        AUC_boot[b] <- NA_real_
      }
      
      # PR-AUC
      PR_boot[b] <- pr_auc_ap(y_b, p_b)
      
      # Accuracy
      ACC_boot[b] <- mean(c_b == y_b)
      
      # Brier
      Brier_boot[b] <- mean((p_b - y_b)^2)
      
      # Calibration
      cal_b <- tryCatch(
        calibration_stats(y_b, p_b),
        error = function(e) list(intercept = NA_real_, slope = NA_real_)
      )
      INT_boot[b]   <- cal_b$intercept
      SLOPE_boot[b] <- cal_b$slope
    }
    
    qfun <- function(x) quantile(x[is.finite(x) & !is.na(x)], probs = c(0.025, 0.975))
    
    if (any(is.finite(AUC_boot)))       AUC_CI         <- qfun(AUC_boot)
    if (any(is.finite(PR_boot)))        PR_AUC_CI      <- qfun(PR_boot)
    if (any(is.finite(ACC_boot)))       ACC_CI         <- qfun(ACC_boot)
    if (any(is.finite(Brier_boot)))     Brier_CI       <- qfun(Brier_boot)
    if (any(is.finite(INT_boot)))       calib_int_CI   <- qfun(INT_boot)
    if (any(is.finite(SLOPE_boot)))     calib_slope_CI <- qfun(SLOPE_boot)
  }
  
  #-------------------------------
  # 10. Return everything
  #-------------------------------
  list(
    call = match.call(),
    formula = formula_obj,
    predictors = predictors,
    stan_fit = fit,
    scale_values = scale_values,
    posterior_summary = summ,
    coef_means = list(
      alpha = alpha_mean,
      beta = beta_means
    ),
    predictions = data.frame(
      row_id = rownames(test),
      observed = y_test,
      predicted_prob = prob_test,
      predicted_class = class_test
    ),
    performance = list(
      AUC = auc_val,
      AUC_CI = AUC_CI,
      PR_AUC = pr_auc_val,
      PR_AUC_CI = PR_AUC_CI,
      accuracy = accuracy,
      accuracy_CI = ACC_CI,
      Brier = brier,
      Brier_CI = Brier_CI,
      calib_intercept = cal$intercept,
      calib_intercept_CI = calib_int_CI,
      calib_slope = cal$slope,
      calib_slope_CI = calib_slope_CI
    ),
    train_index = train_index,
    compiled_model = sm
  )
}

```

Function to check if model is saved
```{r}
run_or_load_model <- function(path, fit_expr, verbose = TRUE) {
  # path: where to save the fitted model (.rds)
  # fit_expr: expression that returns the fitted object (e.g., your fit_bayes_logit_ridge call)
  
  if (file.exists(path)) {
    if (verbose) message("Loading saved model from: ", path)
    fit <- readRDS(path)
  } else {
    if (verbose) message("Fitting model (cache not found): ", path)
    fit <- eval.parent(substitute(fit_expr))
    
    # Ensure directory exists
    dir.create(dirname(path), recursive = TRUE, showWarnings = FALSE)
    
    if (verbose) message("Saving fitted model to: ", path)
    saveRDS(fit, path)
  }
  
  fit
}
```

```{r}
predictors_m1.1 <- c("temperature", "precipitation", "wind", "relative_humidity")#, 
                   #"FFMC", "DMC", "DC", 
                   #"C", "B", "U")
model_path <- file.path(rds_root, "ignition_efficiency_bayes_models/m1.1_bayes_2001_2012_climate.rds")

m1.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              
  predictors = predictors_m1.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))


ggplot() + geom_point(data=m1.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m1.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m2.1 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   "FFMC", "DMC", "DC")#, 
                   #"C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m2.1_bayes_2001_2012_climate.rds")

m2.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m2.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m2.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m2.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m3.1 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   #"FFMC", "DMC", "DC")#, 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m3.1_bayes_2001_2012_climate.rds")

m3.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m3.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m3.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m3.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m4.1 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC")#, 
                   #"C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m4.1_bayes_2001_2012_climate.rds")

m4.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m4.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m4.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m4.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m5.1 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   #"FFMC", "DMC", "DC")#, 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m5.1_bayes_2001_2012_climate.rds")

m5.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m5.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m5.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m5.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m6.1 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m6.1_bayes_2001_2012_climate.rds")

m6.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m6.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m6.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m6.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m7.1 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m7.1_bayes_2001_2012_climate.rds")

m7.1 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m7.1,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m7.1$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m7.1 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)
```

```{r}
model <- c("m1", "m2", "m3", "m4", "m5", "m6", "m7")
training_years <- rep("2001-2012", 7)
approach <- rep("Bayesian", 7)
predictors <- c("climate", "fuel moisture", "vegetation", "climate & fuel moisture", "climate & vegetation", "fuel moisture & vegetation", "climate, fuel moisture & vegetation")

AUC <- c(m1.1$performance$AUC, m2.1$performance$AUC, m3.1$performance$AUC, m4.1$performance$AUC, m5.1$performance$AUC, m6.1$performance$AUC, m7.1$performance$AUC)

PR_AUC <- c(m1.1$performance$PR_AUC, m2.1$performance$PR_AUC, m3.1$performance$PR_AUC, m4.1$performance$PR_AUC, m5.1$performance$PR_AUC, m6.1$performance$PR_AUC, m7.1$performance$PR_AUC)

accuracy <- c(m1.1$performance$accuracy, m2.1$performance$accuracy, m3.1$performance$accuracy, m4.1$performance$accuracy, m5.1$performance$accuracy, m6.1$performance$accuracy, m7.1$performance$accuracy)

Brier <- c(m1.1$performance$Brier, m2.1$performance$Brier, m3.1$performance$Brier, m4.1$performance$Brier, m5.1$performance$Brier, m6.1$performance$Brier, m7.1$performance$Brier)

calib_intercept <- c(m1.1$performance$calib_intercept, m2.1$performance$calib_intercept, m3.1$performance$calib_intercept, m4.1$performance$calib_intercept, m5.1$performance$calib_intercept, m6.1$performance$calib_intercept, m7.1$performance$calib_intercept)

calib_slope <- c(m1.1$performance$calib_slope, m2.1$performance$calib_slope, m3.1$performance$calib_slope, m4.1$performance$calib_slope, m5.1$performance$calib_slope, m6.1$performance$calib_slope, m7.1$performance$calib_slope)

#confidence intervals
AUC_CI_low <- c(m1.1$performance$AUC_CI[1], m2.1$performance$AUC_CI[1], m3.1$performance$AUC_CI[1], m4.1$performance$AUC_CI[1], m5.1$performance$AUC_CI[1], m6.1$performance$AUC_CI[1], m7.1$performance$AUC_CI[1])
AUC_CI_high <- c(m1.1$performance$AUC_CI[2], m2.1$performance$AUC_CI[2], m3.1$performance$AUC_CI[2], m4.1$performance$AUC_CI[2], m5.1$performance$AUC_CI[2], m6.1$performance$AUC_CI[2], m7.1$performance$AUC_CI[2])

PR_AUC_CI_low <- c(m1.1$performance$PR_AUC_CI[1], m2.1$performance$PR_AUC_CI[1], m3.1$performance$PR_AUC_CI[1], m4.1$performance$PR_AUC_CI[1], m5.1$performance$PR_AUC_CI[1], m6.1$performance$PR_AUC_CI[1], m7.1$performance$PR_AUC_CI[1])
PR_AUC_CI_high <- c(m1.1$performance$PR_AUC_CI[2], m2.1$performance$PR_AUC_CI[2], m3.1$performance$PR_AUC_CI[2], m4.1$performance$PR_AUC_CI[2], m5.1$performance$PR_AUC_CI[2], m6.1$performance$PR_AUC_CI[2], m7.1$performance$PR_AUC_CI[2])

Brier_CI_low <- c(m1.1$performance$Brier_CI[1], m2.1$performance$Brier_CI[1], m3.1$performance$Brier_CI[1], m4.1$performance$Brier_CI[1], m5.1$performance$Brier_CI[1], m6.1$performance$Brier_CI[1], m7.1$performance$Brier_CI[1])
Brier_CI_high <- c(m1.1$performance$Brier_CI[2], m2.1$performance$Brier_CI[2], m3.1$performance$Brier_CI[2], m4.1$performance$Brier_CI[2], m5.1$performance$Brier_CI[2], m6.1$performance$Brier_CI[2], m7.1$performance$Brier_CI[2])

calib_intercept_CI_low <- c(m1.1$performance$calib_intercept_CI[1], m2.1$performance$calib_intercept_CI[1], m3.1$performance$calib_intercept_CI[1], m4.1$performance$calib_intercept_CI[1], m5.1$performance$calib_intercept_CI[1], m6.1$performance$calib_intercept_CI[1], m7.1$performance$calib_intercept_CI[1])
calib_intercept_CI_high <- c(m1.1$performance$calib_intercept_CI[2], m2.1$performance$calib_intercept_CI[2], m3.1$performance$calib_intercept_CI[2], m4.1$performance$calib_intercept_CI[2], m5.1$performance$calib_intercept_CI[2], m6.1$performance$calib_intercept_CI[2], m7.1$performance$calib_intercept_CI[2])

calib_slope_CI_low <- c(m1.1$performance$calib_slope_CI[1], m2.1$performance$calib_slope_CI[1], m3.1$performance$calib_slope_CI[1], m4.1$performance$calib_slope_CI[1], m5.1$performance$calib_slope_CI[1], m6.1$performance$calib_slope_CI[1], m7.1$performance$calib_slope_CI[1])
calib_slope_CI_high <- c(m1.1$performance$calib_slope_CI[2], m2.1$performance$calib_slope_CI[2], m3.1$performance$calib_slope_CI[2], m4.1$performance$calib_slope_CI[2], m5.1$performance$calib_slope_CI[2], m6.1$performance$calib_slope_CI[2], m7.1$performance$calib_slope_CI[2])

df_performance_3 <- as.data.frame(cbind(model, training_years, approach, predictors, AUC, PR_AUC, accuracy, Brier, calib_intercept, calib_slope, AUC_CI_low, AUC_CI_high, PR_AUC_CI_low, PR_AUC_CI_high, Brier_CI_low, Brier_CI_high, calib_intercept_CI_low, calib_intercept_CI_high, calib_slope_CI_low, calib_slope_CI_high))

p1 <- ggplot(data=df_performance_3, aes(y=AUC, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p2 <- ggplot(data=df_performance_3, aes(y=accuracy, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Accuracy")

p3 <- ggplot(data=df_performance_3, aes(y=Brier, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Brier")

p1 + p2 + p3
```


#----------------------------
## Data for 2012 - 2018
#----------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2012_2018_10x.csv")
df <- read.csv(drivers_file_path)

df <- na.omit(df) 

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```

```{r}
predictors_m1.2 <- c("temperature", "precipitation", "wind", "relative_humidity")#, 
                   #"FFMC", "DMC", "DC", 
                   #"C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m1.2_bayes_2012_2018_climate.rds")

m1.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m1.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m1.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m1.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m2.2 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   "FFMC", "DMC", "DC")#, 
                   #"C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m2.2_bayes_2012_2018_climate.rds")

m2.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m2.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m2.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m2.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m3.2 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   #"FFMC", "DMC", "DC")#, 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m3.2_bayes_2012_2018_climate.rds")

m3.2 <- run_or_load_model(
  path = model_path,fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m3.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m3.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m3.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m4.2 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC")#, 
                   #"C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m4.2_bayes_2012_2018_climate.rds")

m4.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m4.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m4.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m4.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m5.2 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   #"FFMC", "DMC", "DC")#, 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m5.2_bayes_2012_2018_climate.rds")

m5.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m5.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m5.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m5.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m6.2 <- c(#"temperature", "precipitation", "wind", "relative_humidity")#, 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m6.2_bayes_2012_2018_climate.rds")

m6.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m6.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m6.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m6.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)

#################

predictors_m7.2 <- c("temperature", "precipitation", "wind", "relative_humidity", 
                   "FFMC", "DMC", "DC", 
                   "C", "B", "U")
model_path <- file.path(rds_rood, "ignition_efficiency_bayes_models/m7.2_bayes_2012_2018_climate.rds")

m7.2 <- run_or_load_model(
  path = model_path,
  fit_bayes_logit_ridge(
  data       = df,              # e.g., your 2012–2018 Alaska subset
  predictors = predictors_m7.2,
  response   = "caused_fire",
  sigma_beta = 1,               # prior SD for coefficients (ridge strength)
  sigma_alpha = 5,              # vague prior for intercept
  iter       = 2000,
  chains     = 4
))

ggplot() + geom_point(data=m7.2$predictions, aes(x=as.factor(observed), y=predicted_prob, color=as.factor(predicted_class))) +
  ggtitle("m7.2 Observed vs. Predicted (bayes approach)") +
  xlab("Observed") + ylab("Predicted Probability") +
  labs(color = "Predicted Class") +
  ylim(0,1)
```

```{r}
model <- c("m1", "m2", "m3", "m4", "m5", "m6", "m7")
training_years <- rep("2012-2018", 7)
approach <- rep("Bayesian", 7)
predictors <- c("climate", "fuel moisture", "vegetation", "climate & fuel moisture", "climate & vegetation", "fuel moisture & vegetation", "climate, fuel moisture & vegetation")

AUC <- c(m1.2$performance$AUC, m2.2$performance$AUC, m3.2$performance$AUC, m4.2$performance$AUC, m5.2$performance$AUC, m6.2$performance$AUC, m7.2$performance$AUC)

PR_AUC <- c(m1.2$performance$PR_AUC, m2.2$performance$PR_AUC, m3.2$performance$PR_AUC, m4.2$performance$PR_AUC, m5.2$performance$PR_AUC, m6.2$performance$PR_AUC, m7.2$performance$PR_AUC)

accuracy <- c(m1.2$performance$accuracy, m2.2$performance$accuracy, m3.2$performance$accuracy, m4.2$performance$accuracy, m5.2$performance$accuracy, m6.2$performance$accuracy, m7.2$performance$accuracy)

Brier <- c(m1.2$performance$Brier, m2.2$performance$Brier, m3.2$performance$Brier, m4.2$performance$Brier, m5.2$performance$Brier, m6.2$performance$Brier, m7.2$performance$Brier)

calib_intercept <- c(m1.2$performance$calib_intercept, m2.2$performance$calib_intercept, m3.2$performance$calib_intercept, m4.2$performance$calib_intercept, m5.2$performance$calib_intercept, m6.2$performance$calib_intercept, m7.2$performance$calib_intercept)

calib_slope <- c(m1.2$performance$calib_slope, m2.2$performance$calib_slope, m3.2$performance$calib_slope, m4.2$performance$calib_slope, m5.2$performance$calib_slope, m6.2$performance$calib_slope, m7.2$performance$calib_slope)

#confidence intervals
AUC_CI_low <- c(m1.2$performance$AUC_CI[1], m2.2$performance$AUC_CI[1], m3.2$performance$AUC_CI[1], m4.2$performance$AUC_CI[1], m5.2$performance$AUC_CI[1], m6.2$performance$AUC_CI[1], m7.2$performance$AUC_CI[1])
AUC_CI_high <- c(m1.2$performance$AUC_CI[2], m2.2$performance$AUC_CI[2], m3.2$performance$AUC_CI[2], m4.2$performance$AUC_CI[2], m5.2$performance$AUC_CI[2], m6.2$performance$AUC_CI[2], m7.2$performance$AUC_CI[2])

PR_AUC_CI_low <- c(m1.2$performance$PR_AUC_CI[1], m2.2$performance$PR_AUC_CI[1], m3.2$performance$PR_AUC_CI[1], m4.2$performance$PR_AUC_CI[1], m5.2$performance$PR_AUC_CI[1], m6.2$performance$PR_AUC_CI[1], m7.2$performance$PR_AUC_CI[1])
PR_AUC_CI_high <- c(m1.2$performance$PR_AUC_CI[2], m2.2$performance$PR_AUC_CI[2], m3.2$performance$PR_AUC_CI[2], m4.2$performance$PR_AUC_CI[2], m5.2$performance$PR_AUC_CI[2], m6.2$performance$PR_AUC_CI[2], m7.2$performance$PR_AUC_CI[2])

Brier_CI_low <- c(m1.2$performance$Brier_CI[1], m2.2$performance$Brier_CI[1], m3.2$performance$Brier_CI[1], m4.2$performance$Brier_CI[1], m5.2$performance$Brier_CI[1], m6.2$performance$Brier_CI[1], m7.2$performance$Brier_CI[1])
Brier_CI_high <- c(m1.2$performance$Brier_CI[2], m2.2$performance$Brier_CI[2], m3.2$performance$Brier_CI[2], m4.2$performance$Brier_CI[2], m5.2$performance$Brier_CI[2], m6.2$performance$Brier_CI[2], m7.2$performance$Brier_CI[2])

calib_intercept_CI_low <- c(m1.2$performance$calib_intercept_CI[1], m2.2$performance$calib_intercept_CI[1], m3.2$performance$calib_intercept_CI[1], m4.2$performance$calib_intercept_CI[1], m5.2$performance$calib_intercept_CI[1], m6.2$performance$calib_intercept_CI[1], m7.2$performance$calib_intercept_CI[1])
calib_intercept_CI_high <- c(m1.2$performance$calib_intercept_CI[2], m2.2$performance$calib_intercept_CI[2], m3.2$performance$calib_intercept_CI[2], m4.2$performance$calib_intercept_CI[2], m5.2$performance$calib_intercept_CI[2], m6.2$performance$calib_intercept_CI[2], m7.2$performance$calib_intercept_CI[2])

calib_slope_CI_low <- c(m1.2$performance$calib_slope_CI[1], m2.2$performance$calib_slope_CI[1], m3.2$performance$calib_slope_CI[1], m4.2$performance$calib_slope_CI[1], m5.2$performance$calib_slope_CI[1], m6.2$performance$calib_slope_CI[1], m7.2$performance$calib_slope_CI[1])
calib_slope_CI_high <- c(m1.2$performance$calib_slope_CI[2], m2.2$performance$calib_slope_CI[2], m3.2$performance$calib_slope_CI[2], m4.2$performance$calib_slope_CI[2], m5.2$performance$calib_slope_CI[2], m6.2$performance$calib_slope_CI[2], m7.2$performance$calib_slope_CI[2])

df_performance_4 <- as.data.frame(cbind(model, training_years, approach, predictors, AUC, PR_AUC, accuracy, Brier, calib_intercept, calib_slope, AUC_CI_low, AUC_CI_high, PR_AUC_CI_low, PR_AUC_CI_high, Brier_CI_low, Brier_CI_high,  calib_intercept_CI_low, calib_intercept_CI_high, calib_slope_CI_low, calib_slope_CI_high))

p1 <- ggplot(data=df_performance_4, aes(y=AUC, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Area Under the Curve (AUC)")

p2 <- ggplot(data=df_performance_4, aes(y=accuracy, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Accuracy")

p3 <- ggplot(data=df_performance_4, aes(y=Brier, x=model, fill=predictors)) + geom_bar(stat="identity", position=position_dodge()) + ggtitle("Brier")

p1 + p2 + p3
```


```{r}
performance_df <- as.data.frame(rbind(df_performance_1, df_performance_2, df_performance_3, df_performance_4))
performance_df$AUC <- as.numeric(performance_df$AUC)
performance_df$PR_AUC <- as.numeric(performance_df$PR_AUC)
performance_df$Brier <- as.numeric(performance_df$Brier)
performance_df$calib_intercept <- as.numeric(performance_df$calib_intercept)
performance_df$calib_slope <- as.numeric(performance_df$calib_slope)

performance_df$AUC_CI_low <- as.numeric(performance_df$AUC_CI_low)
performance_df$AUC_CI_high <- as.numeric(performance_df$AUC_CI_high)
performance_df$PR_AUC_CI_low <- as.numeric(performance_df$PR_AUC_CI_low)
performance_df$PR_AUC_CI_high <- as.numeric(performance_df$PR_AUC_CI_high)
performance_df$Brier_CI_low <- as.numeric(performance_df$Brier_CI_low)
performance_df$Brier_CI_high <- as.numeric(performance_df$Brier_CI_high)
performance_df$calib_intercept_CI_low <- as.numeric(performance_df$calib_intercept_CI_low)
performance_df$calib_intercept_CI_high <- as.numeric(performance_df$calib_intercept_CI_high)
performance_df$calib_slope_CI_low <- as.numeric(performance_df$calib_slope_CI_low)
performance_df$calib_slope_CI_high <- as.numeric(performance_df$calib_slope_CI_high)

table_file_path <- file.path(table_root, "ignition_efficiency_model_performance.csv")
write.csv(performance_df, table_file_path)
```

