---
title: "Random Forest"
author: "Charlotte Uden"
date: "2025-08-29"
output: html_document
---

```{r}
library(randomForest)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]
```

---------------------------------------------------
Run random forest for 2002-2011 ignition efficiency
---------------------------------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2002_2011_10x.csv")
df <- read.csv(drivers_file_path)[,2:13]
df <- na.omit(df) 

df <- df %>%
  drop_na()

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```


```{r}
df <- df[c("caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]

#check for correlation between variables
#draw random sample of rows to plot 
plot(df)

#apply random forest:
rf <-randomForest(caused_fire~.,data=df, ntree=500) 
print(rf)

#mtry: Number of random variables used in each tree
# - reducing mtry reduces correlation and strength. 
# - the oob error rate tells you what value of mtry is optimal

mtry <- tuneRF(df[-1],df$caused_fire, ntreeTry=500,
               stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
#looking for mtry for the lowest oob error rate 
print(mtry)
print(best.m) 

#apply best mtry to model
rf <-randomForest(caused_fire~.,data=df, mtry=best.m, importance=TRUE,ntree=500)
print(rf)

#Evaluate variable importance
importance(rf)
varImpPlot(rf)


```
# Mean Decrease Accuracy (%IncMSE):
- remove variables from the model and asses which have the greatest impact on model accuracy.
- Higher the value of mean decrease accuracy or mean decrease gini score ,
       higher the importance of the variable in the model.
- tells you effect of randomly shuffling values of a variable on increasing mse of model predictions
- a higher number indicates that the variable is more important

# Mean Decrease Gini or Increase in Node Purity (IncNodePurity):
- expresses the change in the homogeneity of the of the groups created by the trees
     (using the Gini coefficient as a measure).
- What is expressed is the decrease in said purity if a particular variable
      has no information. If a variable has no information to begin with,
      the decrase would be zero. (from: https://www.reagro.org/methods/statistical/randomforest.html)

Save random forest results forAppendix
```{r}
rf_results <- importance(rf)
rf_results <- data.frame(Variable = rownames(rf_results), 
                     IncMSE = rf_results[, "%IncMSE"],
                     IncNodePurity = rf_results[, "IncNodePurity"]) 
# Export
table_file_path <- file.path(table_root, "lightning_random_forest_results_2002_2011.csv")
write.csv(rf_results, table_file_path, row.names = FALSE)
```


---------------------------------------------------
Run random forest for 2012-2018 ignition efficiency
---------------------------------------------------

```{r}
drivers_file_path <- file.path(drivers_root, "ignition_efficiency_drivers_2012_2018_10x.csv")
df <- read.csv(drivers_file_path)[,2:13]
df <- na.omit(df) 

df <- df %>%
  drop_na()

#note that lat.1 and lon.1, and date.1 are the locations and date of the era5 grid points. 
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d")
df <- df[ , c("lat","lon","date","caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]
str(df)
```


```{r}
df <- df[c("caused_fire","temperature","precipitation","wind","relative_humidity","FFMC","DMC","DC", "C", "B", "U")]

#check for correlation between variables
#draw random sample of rows to plot 
plot(df)

#apply random forest:
rf <-randomForest(caused_fire~.,data=df, ntree=500) 
print(rf)

#mtry: Number of random variables used in each tree
# - reducing mtry reduces correlation and strength. 
# - the oob error rate tells you what value of mtry is optimal

mtry <- tuneRF(df[-1],df$caused_fire, ntreeTry=500,
               stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]
#looking for mtry for the lowest oob error rate 
print(mtry)
print(best.m) 

#apply best mtry to model
rf <-randomForest(caused_fire~.,data=df, mtry=best.m, importance=TRUE,ntree=500)
print(rf)

#Evaluate variable importance
importance(rf)
varImpPlot(rf)


```

Save random forest results for Appendix  
```{r}
rf_results <- importance(rf)
rf_results <- data.frame(Variable = rownames(rf_results), 
                     IncMSE = rf_results[, "%IncMSE"],
                     IncNodePurity = rf_results[, "IncNodePurity"]) 
# Export
table_file_path <- file.path(table_root, "lightning_random_forest_results_2012_2018.csv")
write.csv(rf_results, table_file_path, row.names = FALSE)
```




