---
title: "Lightning Prediction, Alaska"
author: "Charlotte Uden"
date: "2025-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
library(rstan)
library(dplyr)
library(stats)
library(caret)
library(tidyr)
library(patchwork)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths (ie line 52)
```

#---------------------
2002-2011
#---------------------

```{r}
drivers_file_path <- file.path(drivers_root, "lightning_prediction_drivers_2002_2011.csv")
df <- read.csv(drivers_file_path)[,2:13]

colnames(df) <- c("year", "lon", "lat", "strikes", "tair", "precip", "wind", "swr", "sp", "rh", "cape", "cxp")
head(df)
```

2. Split into training and testing data
```{r, echo=FALSE}
## 80% of the sample size
sample_size <- floor(0.8 * nrow(df))

## set the seed to make partition reproducible
set.seed(123)
train_index <- sample(seq_len(nrow(df)), size = sample_size)

train <- df[train_index, ]
test <- df[-train_index, ]

```


#Standardize the data
```{r}
mu_precip <- mean(train$precip)
sigma_precip <- sd(train$precip)
mu_tair <- mean(train$tair)
sigma_tair <- sd(train$tair)
mu_wind <- mean(train$wind)
sigma_wind <- sd(train$wind)
mu_swr <- mean(train$swr)
sigma_swr <- sd(train$swr)
mu_sp <- mean(train$sp)
sigma_sp <- sd(train$sp)
mu_rh <- mean(train$rh)
sigma_rh <- sd(train$rh)
mu_cape <- mean(train$cape)
sigma_cape <- sd(train$cape)
mu_cxp <- mean(train$cxp)
sigma_cxp <- sd(train$cxp)
#-----

train <- train %>%
  mutate(precip = (precip-mu_precip)/sigma_precip) %>%
  mutate(tair = (tair-mu_tair)/sigma_tair) %>%
  mutate(wind = (wind-mu_wind)/sigma_wind) %>%
  mutate(swr = (swr-mu_swr)/sigma_swr) %>%
  mutate(sp = (sp-mu_sp)/sigma_sp) %>%
  mutate(rh = (rh-mu_rh)/sigma_rh) %>%
  mutate(cape = (cape-mu_cape)/sigma_cape) %>%
  mutate(cxp = (cxp-mu_cxp)/sigma_cxp)



train <- train[,c("strikes","precip","tair", "wind", "swr", "sp", "rh", "cape", "cxp")]
head(train)


test <- test %>%
  mutate(precip = (precip-mu_precip)/sigma_precip) %>%
  mutate(tair = (tair-mu_tair)/sigma_tair) %>%
  mutate(wind = (wind-mu_wind)/sigma_wind) %>%
  mutate(swr = (swr-mu_swr)/sigma_swr) %>%
  mutate(sp = (sp-mu_sp)/sigma_sp) %>%
  mutate(rh = (rh-mu_rh)/sigma_rh) %>%
  mutate(cape = (cape-mu_cape)/sigma_cape) %>%
  mutate(cxp = (cxp-mu_cxp)/sigma_cxp) %>%
  mutate(lon = lon) %>%
  mutate(lat = lat)

test <- test[,c("lon", "lat", "year", "strikes", "precip", "tair", "wind", "swr", "sp", "rh", "cape", "cxp")]

plot(test)

hist(test$strikes)

```
The gamma distribution cannot handle 0 values. the monthly summaries do have 0's, so we need to add a very small value so that the gamma distribution can handle the data:
```{r}
epsilon <- 1e-4  # small value
train$strikes<- train$strikes + epsilon
```

Functions for calculating performance metrics
```{r}
# ============================================================
# Bootstrap CI versions of each performance metric
# Each function returns: estimate, ci_low, ci_high
# Percentile CI: 2.5% and 97.5% from n_boot bootstrap samples
# Resampling is done on the TEST SET rows / paired vectors
# ============================================================

# ---- helpers (small + boring, used by all) ----
.rmse <- function(obs, pred) {
  ok <- is.finite(obs) & is.finite(pred)
  obs <- obs[ok]; pred <- pred[ok]
  if (length(obs) == 0) return(NA_real_)
  sqrt(mean((obs - pred)^2))
}

.safe_cor <- function(x, y) {
  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]
  if (length(x) < 2) return(NA_real_)
  if (sd(x) == 0 || sd(y) == 0) return(NA_real_)
  cor(x, y, use = "complete.obs")
}

.ci_from_boot <- function(boot_vals, conf = 0.95) {
  alpha <- (1 - conf) / 2
  boot_vals <- boot_vals[is.finite(boot_vals)]
  if (length(boot_vals) == 0) {
    return(list(ci_low = NA_real_, ci_high = NA_real_))
  }
  list(
    ci_low  = as.numeric(stats::quantile(boot_vals, probs = alpha, na.rm = TRUE)),
    ci_high = as.numeric(stats::quantile(boot_vals, probs = 1 - alpha, na.rm = TRUE))
  )
}

# ============================================================
# 1) Normalized RMSE (normalized by SD of observed)
# ============================================================
metric_nrmse_ci <- function(obs, pred, n_boot = 1000, conf = 0.95, seed = 123) {
  ok <- is.finite(obs) & is.finite(pred)
  obs <- obs[ok]; pred <- pred[ok]
  N <- length(obs)

  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  denom <- sd(obs)
  est <- if (!is.finite(denom) || denom == 0) NA_real_ else .rmse(obs, pred) / denom

  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    o <- obs[idx]; p <- pred[idx]
    d <- sd(o)
    if (!is.finite(d) || d == 0) return(NA_real_)
    .rmse(o, p) / d
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# 2) Correlation coefficient (Pearson)
# ============================================================
metric_cor_ci <- function(obs, pred, n_boot = 1000, conf = 0.95, seed = 123) {
  ok <- is.finite(obs) & is.finite(pred)
  obs <- obs[ok]; pred <- pred[ok]
  N <- length(obs)

  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  est <- .safe_cor(obs, pred)

  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    .safe_cor(obs[idx], pred[idx])
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# 3) Skill score (Perkins S-score)
# ============================================================
metric_sscore_ci <- function(obs, pred, nbins = 15,
                             n_boot = 1000, conf = 0.95, seed = 123) {
  ok <- is.finite(obs) & is.finite(pred)
  obs <- obs[ok]; pred <- pred[ok]
  N <- length(obs)

  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  calculate_sscore <- function(obs, pred, nbins = 15) {
    min_val <- min(c(obs, pred))
    max_val <- max(c(obs, pred))
    if (!is.finite(min_val) || !is.finite(max_val) || min_val == max_val) return(NA_real_)

    breaks <- seq(min_val, max_val, length.out = nbins + 1)
    hist_obs  <- hist(obs,  breaks = breaks, plot = FALSE)
    hist_pred <- hist(pred, breaks = breaks, plot = FALSE)

    prob_obs  <- hist_obs$counts  / sum(hist_obs$counts)
    prob_pred <- hist_pred$counts / sum(hist_pred$counts)

    sum(pmin(prob_obs, prob_pred))
  }

  est <- calculate_sscore(obs, pred, nbins = nbins)

  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    calculate_sscore(obs[idx], pred[idx], nbins = nbins)
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# 4) Spatial correlation
#    (correlation of gridcell means across space)
#    test must contain lon, lat, and strikes (obs)
# ============================================================
metric_spatial_cor_ci <- function(test, y_pred,
                                  n_boot = 1000, conf = 0.95, seed = 123) {
  test <- dplyr::as_tibble(test)
  test$pred <- y_pred

  test <- test %>% dplyr::filter(is.finite(strikes), is.finite(pred))
  N <- nrow(test)
  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  # point estimate on full test set
  spatial_means <- test %>%
    dplyr::group_by(lon, lat) %>%
    dplyr::summarise(
      obs_mean  = mean(strikes, na.rm = TRUE),
      pred_mean = mean(pred,    na.rm = TRUE),
      .groups = "drop"
    )

  est <- .safe_cor(spatial_means$obs_mean, spatial_means$pred_mean)

  # bootstrap resampling rows of the test set
  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    tb <- test[idx, , drop = FALSE]

    sm <- tb %>%
      dplyr::group_by(lon, lat) %>%
      dplyr::summarise(
        obs_mean  = mean(strikes, na.rm = TRUE),
        pred_mean = mean(pred,    na.rm = TRUE),
        .groups = "drop"
      )

    .safe_cor(sm$obs_mean, sm$pred_mean)
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# 5) Anomaly Correlation Coefficient (ACC)
#    computed from annual means, then anomalies, then correlation
# ============================================================
metric_acc_ci <- function(year, obs, pred, n_boot = 1000, conf = 0.95, seed = 123) {
  df <- dplyr::tibble(year = year, obs = obs, pred = pred) %>%
    dplyr::filter(is.finite(year), is.finite(obs), is.finite(pred))

  N <- nrow(df)
  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  calc_acc_core <- function(df_in) {
    annual <- df_in %>%
      dplyr::group_by(year) %>%
      dplyr::summarise(
        obs  = mean(obs,  na.rm = TRUE),
        pred = mean(pred, na.rm = TRUE),
        .groups = "drop"
      )
    obs_anom  <- annual$obs  - mean(annual$obs,  na.rm = TRUE)
    pred_anom <- annual$pred - mean(annual$pred, na.rm = TRUE)
    .safe_cor(obs_anom, pred_anom)
  }

  est <- calc_acc_core(df)

  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    calc_acc_core(df[idx, , drop = FALSE])
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# 6) Normalized RMSE of anomalies (nRMSE_anom)
#    = RMSE(annual anomalies) / SD(observed annual anomalies)
# ============================================================
metric_nrmse_anom_ci <- function(year, obs, pred, n_boot = 1000, conf = 0.95, seed = 123) {
  df <- dplyr::tibble(year = year, obs = obs, pred = pred) %>%
    dplyr::filter(is.finite(year), is.finite(obs), is.finite(pred))

  N <- nrow(df)
  if (N == 0) return(list(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))

  calc_nrmse_anom_core <- function(df_in) {
    annual <- df_in %>%
      dplyr::group_by(year) %>%
      dplyr::summarise(
        obs  = mean(obs,  na.rm = TRUE),
        pred = mean(pred, na.rm = TRUE),
        .groups = "drop"
      )
    obs_anom  <- annual$obs  - mean(annual$obs,  na.rm = TRUE)
    pred_anom <- annual$pred - mean(annual$pred, na.rm = TRUE)

    denom <- sd(obs_anom, na.rm = TRUE)
    if (!is.finite(denom) || denom == 0) return(NA_real_)
    sqrt(mean((obs_anom - pred_anom)^2, na.rm = TRUE)) / denom
  }

  est <- calc_nrmse_anom_core(df)

  set.seed(seed)
  boot <- replicate(n_boot, {
    idx <- sample.int(N, size = N, replace = TRUE)
    calc_nrmse_anom_core(df[idx, , drop = FALSE])
  })

  ci <- .ci_from_boot(boot, conf = conf)
  list(estimate = est, ci_low = ci$ci_low, ci_high = ci$ci_high)
}

# ============================================================
# Example calls
# ============================================================
# nrmse_res  <- metric_nrmse_ci(test$strikes, y_pred)
# cor_res    <- metric_cor_ci(test$strikes, y_pred)
# sscore_res <- metric_sscore_ci(test$strikes, y_pred, nbins = 15)
# spcor_res  <- metric_spatial_cor_ci(test, y_pred)
# acc_res    <- metric_acc_ci(test$year, test$strikes, y_pred)
# nrmseA_res <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred)

```

Linear Model
```{r}
#cape
N1 <- lm(strikes ~ cape,
                 data = train)

a <- N1$coefficients[1]
b <- N1$coefficients[2]


N1_coefficients <- c(a, b)

y_pred_N1 <- a + b*test$cape 

nrmse_res_N1  <- metric_nrmse_ci(test$strikes, y_pred_N1)
cor_res_N1   <- metric_cor_ci(test$strikes, y_pred_N1)
sscore_res_N1 <- metric_sscore_ci(test$strikes, y_pred_N1, nbins = 15)
spcor_res_N1  <- metric_spatial_cor_ci(test, y_pred_N1)
acc_res_N1  <- metric_acc_ci(test$year, test$strikes, y_pred_N1)
nrmseA_res_N1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N1)

#CAPE x Precip (cxp)
N2 <- lm(strikes ~ cxp,
                 data = train)

a <- N2$coefficients[1]
b <- N2$coefficients[2]

N2_coefficients <- c(a, b)

y_pred_N2 <- a + b*test$cxp

nrmse_res_N2  <- metric_nrmse_ci(test$strikes, y_pred_N2)
cor_res_N2    <- metric_cor_ci(test$strikes, y_pred_N2)
sscore_res_N2 <- metric_sscore_ci(test$strikes, y_pred_N2, nbins = 15)
spcor_res_N2  <- metric_spatial_cor_ci(test, y_pred_N2)
acc_res_N2  <- metric_acc_ci(test$year, test$strikes, y_pred_N2)
nrmseA_res_N2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N2)

#swr + tair + rh + wind + precip +sp
N13 <- lm(strikes ~ swr + tair + rh + wind + precip +sp,
                 data = train)

a <- N13$coefficients[1]
b <- N13$coefficients[2]
c <- N13$coefficients[3]
d <- N13$coefficients[4]
e <- N13$coefficients[5]
f <- N13$coefficients[6]
g <- N13$coefficients[7]

N13_coefficients <- c(a, b, c, d, e, f, g)

y_pred_N13 <- a + b*test$swr + c*test$tair + d*test$rh +e*test$wind + f*test$precip + g*test$sp

nrmse_res_N13  <- metric_nrmse_ci(test$strikes, y_pred_N13)
cor_res_N13    <- metric_cor_ci(test$strikes, y_pred_N13)
sscore_res_N13 <- metric_sscore_ci(test$strikes, y_pred_N13, nbins = 15)
spcor_res_N13  <- metric_spatial_cor_ci(test, y_pred_N13)
acc_res_N13  <- metric_acc_ci(test$year, test$strikes, y_pred_N13)
nrmseA_res_N13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N13)

```

Gamma GLM
```{r}
#cape
G1 <- glm(strikes ~ cape,
                 data = train,
                 family = Gamma(link =log))

a <- G1$coefficients[1]
b <- G1$coefficients[2]

G1_coefficients <- c(a, b)

y_pred_G1 <- exp(a + b*test$cape)

nrmse_res_G1  <- metric_nrmse_ci(test$strikes, y_pred_G1)
cor_res_G1    <- metric_cor_ci(test$strikes, y_pred_G1)
sscore_res_G1 <- metric_sscore_ci(test$strikes, y_pred_G1, nbins = 15)
spcor_res_G1  <- metric_spatial_cor_ci(test, y_pred_G1)
acc_res_G1  <- metric_acc_ci(test$year, test$strikes, y_pred_G1)
nrmseA_res_G1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G1)

#CAPE x precip (cxp)
G2 <- glm(strikes ~ cxp,
                 data = train,
                 family = Gamma(link =log))

a <- G2$coefficients[1]
b <- G2$coefficients[2]

G2_coefficients <- c(a, b)

y_pred_G2 <- exp(a + b*test$cape)

nrmse_res_G2  <- metric_nrmse_ci(test$strikes, y_pred_G2)
cor_res_G2    <- metric_cor_ci(test$strikes, y_pred_G2)
sscore_res_G2 <- metric_sscore_ci(test$strikes, y_pred_G2, nbins = 15)
spcor_res_G2  <- metric_spatial_cor_ci(test, y_pred_G2)
acc_res_G2  <- metric_acc_ci(test$year, test$strikes, y_pred_G2)
nrmseA_res_G2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G2)

#swr + tair + rh + wind + precip +sp
G13 <- glm(strikes ~ swr + tair + rh + wind + precip + sp,
                 data = train,
                 family = Gamma(link =log))

a <- G13$coefficients[1]
b <- G13$coefficients[2]
c <- G13$coefficients[3]
d <- G13$coefficients[4]
e <- G13$coefficients[5]
f <- G13$coefficients[6]
g <- G13$coefficients[7]

G13_coefficients <- c(a, b, c, d, e, f, g)

y_pred_G13 <- exp(a + b*test$swr + c*test$tair + d*test$rh +e*test$wind + f*test$precip + g*test$sp)

nrmse_res_G13  <- metric_nrmse_ci(test$strikes, y_pred_G13)
cor_res_G13    <- metric_cor_ci(test$strikes, y_pred_G13)
sscore_res_G13 <- metric_sscore_ci(test$strikes, y_pred_G13, nbins = 15)
spcor_res_G13  <- metric_spatial_cor_ci(test, y_pred_G13)
acc_res_G13  <- metric_acc_ci(test$year, test$strikes, y_pred_G13)
nrmseA_res_G13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G13)
```

# B1: CAPE

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: CAPE
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * CAPE,
    beta <- a_beta + b_beta * CAPE,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  CAPE = train$cape
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1,
       a_beta = 1, b_beta = 0.1)
}

if (!file.exists(file.path(rds_root, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2001_2012.rds"))) {
  B1 <- stan(file = "gamma_cape.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B1, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2001_2012.rds"))
    } else {
      B1 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2001_2012.rds"))
    }

```


```{r}

#print(B1, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B1)

parameter_stats <- summary(B1)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]

a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]

B1_alpha_coefficients <- c(a_alpha, b_alpha)

B1_beta_coefficients <- c(a_beta, b_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B1)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha

a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$cape)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$cape[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$cape[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  cape = test$cape,            #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$cape, predictions_df$sampled_pred, color=test$sp, main="Lightning Predictions")
plot(test$cape, test$strikes, color=test$sp, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B1 <- predictions_df$sampled_pred
```

```{r}
nrmse_res_B1  <- metric_nrmse_ci(test$strikes, y_pred_B1)
cor_res_B1    <- metric_cor_ci(test$strikes, y_pred_B1)
sscore_res_B1 <- metric_sscore_ci(test$strikes, y_pred_B1, nbins = 15)
spcor_res_B1  <- metric_spatial_cor_ci(test, y_pred_B1)
acc_res_B1  <- metric_acc_ci(test$year, test$strikes, y_pred_B1)
nrmseA_res_B1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B1)
```

# B2: CAPE x Precipitation (cxp)

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: shortwave radiation,  temperature, relative humidity, wind, precipitation, and surface pressure
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * CXP,
    beta <- a_beta + b_beta * CXP,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  CXP = train$cxp
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1,
       a_beta = 1, b_beta = 0.1)
}

if (!file.exists(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2001_2012.rds"))) {
  B2 <- stan(file = "gamma_cxp.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B2, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2001_2012.rds"))
    } else {
      B2 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2001_2012.rds"))
    }

```


```{r}

#print(B2, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B2)

parameter_stats <- summary(B2)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]

a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]

B2_alpha_coefficients <- c(a_alpha, b_alpha)

B2_beta_coefficients <- c(a_beta, b_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B2)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha

a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$cxp)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$cxp[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$cxp[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  cxp = test$cxp,           #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$cxp, predictions_df$sampled_pred, main="Lightning Predictions")
plot(test$cxp, test$strikes, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B2 <- predictions_df$sampled_pred
```

```{r}
nrmse_res_B2  <- metric_nrmse_ci(test$strikes, y_pred_B2)
cor_res_B2    <- metric_cor_ci(test$strikes, y_pred_B2)
sscore_res_B2 <- metric_sscore_ci(test$strikes, y_pred_B2, nbins = 15)
spcor_res_B2  <- metric_spatial_cor_ci(test, y_pred_B2)
acc_res_B2  <- metric_acc_ci(test$year, test$strikes, y_pred_B2)
nrmseA_res_B2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B2)
```

# B13: Short Wave Radiation + Temperature + Precipitation + Wind + Precipitation + Surface Pressure

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: shortwave radiation,  temperature, relative humidity, wind, precipitation, and surface pressure
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * SWR + c_alpha * T + d_alpha * RH + e_alpha * W + f_alpha * P + g_alpha * SP,
    beta <- a_beta + b_beta * SWR + c_beta * T + d_beta * RH + e_beta * W + f_beta * P + g_beta * SP,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  SWR = train$swr,
  T = train$tair,
  RH = train$rh,
  W = train$wind,
  P = train$precip,
  SP = train$sp
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1, c_alpha = 0.1, d_alpha = 0.1, e_alpha = 0.1, f_alpha = 0.1, g_alpha = 0.1,
       a_beta = 1, b_beta = 0.1, c_beta = 0.1, d_beta = 0.1, e_beta = 0.1, f_beta = 0.1, g_beta = 0.1)
}

if (!file.exists(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2001_2012.rds"))) {
  B13 <- stan(file = "gamma_swr_tair_rh_wind_precip_sp.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B13, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2001_2012.rds"))
    } else {
      B13 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2001_2012.rds"))
    }

```


```{r}

#print(B13, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B13)

parameter_stats <- summary(B13)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]
c_alpha <- summary_table["c_alpha", "mean"]
d_alpha <- summary_table["d_alpha", "mean"]
e_alpha <- summary_table["e_alpha", "mean"]
f_alpha <- summary_table["f_alpha", "mean"]
g_alpha <- summary_table["g_alpha", "mean"]
a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]
c_beta <- summary_table["c_beta", "mean"]
d_beta <- summary_table["d_beta", "mean"]
e_beta <- summary_table["e_beta", "mean"]
f_beta <- summary_table["f_beta", "mean"]
g_beta <- summary_table["g_beta", "mean"]

B13_alpha_coefficients <- c(a_alpha, b_alpha, c_alpha, d_alpha, e_alpha, f_alpha, g_alpha)

B13_beta_coefficients <- c(a_beta, b_beta, c_beta, d_beta, e_beta, f_beta, g_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B13)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha
c_alpha_samples <- posterior_samples$c_alpha
d_alpha_samples <- posterior_samples$d_alpha
e_alpha_samples <- posterior_samples$e_alpha
f_alpha_samples <- posterior_samples$f_alpha
g_alpha_samples <- posterior_samples$g_alpha
a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta
c_beta_samples <- posterior_samples$c_beta
d_beta_samples <- posterior_samples$d_beta
e_beta_samples <- posterior_samples$e_beta
f_beta_samples <- posterior_samples$f_beta
g_beta_samples <- posterior_samples$g_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$swr)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$swr[j] + c_alpha_samples[i] * test$tair[j] + d_alpha_samples[i] * test$rh[j] + e_alpha_samples[i] * test$wind[j] + f_alpha_samples[i] * test$precip[j] + g_alpha_samples[i] * test$sp[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$swr[j] + c_alpha_samples[i] * test$tair[j] + d_alpha_samples[i] * test$rh[j] + e_alpha_samples[i] * test$wind[j] + f_alpha_samples[i] * test$precip[j] + g_alpha_samples[i] * test$sp[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  swr = test$swr,            #Climate predictor
  tair = test$tair,            #Climate predictor
  rh = test$rh,            #Climate predictor
  wind = test$wind,            #Climate predictor
  precip = test$precip,            #Climate predictor
  sp = test$sp,            #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$swr, predictions_df$sampled_pred, color=test$sp, main="Lightning Predictions")
plot(test$swr, test$strikes, color=test$sp, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B13 <- predictions_df$sampled_pred
```

```{r}
nrmse_res_B13  <- metric_nrmse_ci(test$strikes, y_pred_B13)
cor_res_B13    <- metric_cor_ci(test$strikes, y_pred_B13)
sscore_res_B13 <- metric_sscore_ci(test$strikes, y_pred_B13, nbins = 15)
spcor_res_B13  <- metric_spatial_cor_ci(test, y_pred_B13)
acc_res_B13  <- metric_acc_ci(test$year, test$strikes, y_pred_B13)
nrmseA_res_B13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B13)
```

Dataframe of Performance Metrics

```{r}

nrmse <- c(nrmse_res_N1$estimate, nrmse_res_G1$estimate, nrmse_res_B1$estimate, 
           nrmse_res_N2$estimate, nrmse_res_G2$estimate, nrmse_res_B2$estimate, 
           nrmse_res_N13$estimate, nrmse_res_G13$estimate, nrmse_res_B13$estimate)
cor <- c(cor_res_N1$estimate, cor_res_G1$estimate, cor_res_B1$estimate,
         cor_res_N2$estimate, cor_res_G2$estimate, cor_res_B2$estimate,
         cor_res_N13$estimate, cor_res_G13$estimate, cor_res_B13$estimate)
sscore <- c(sscore_res_N1$estimate, sscore_res_G1$estimate, sscore_res_B1$estimate,
            sscore_res_N2$estimate, sscore_res_G2$estimate, sscore_res_B2$estimate,
            sscore_res_N13$estimate, sscore_res_G13$estimate, sscore_res_B13$estimate)
spcor <- c(spcor_res_N1$estimate, spcor_res_G1$estimate, spcor_res_B1$estimate,
           spcor_res_N2$estimate, spcor_res_G2$estimate, spcor_res_B2$estimate,
           spcor_res_N13$estimate, spcor_res_G13$estimate, spcor_res_B13$estimate)
acc <- c(acc_res_N1$estimate, acc_res_G1$estimate, acc_res_B1$estimate, 
         acc_res_N2$estimate, acc_res_G2$estimate, acc_res_B2$estimate,
         acc_res_N13$estimate, acc_res_G13$estimate, acc_res_B13$estimate)
nrmseA <- c(nrmseA_res_N1$estimate, nrmseA_res_G1$estimate, nrmseA_res_B1$estimate,
            nrmseA_res_N2$estimate, nrmseA_res_G2$estimate, nrmseA_res_B2$estimate,
            nrmseA_res_N13$estimate, nrmseA_res_G13$estimate, nrmseA_res_B13$estimate)

nrmse_ci_low <- c(nrmse_res_N1$ci_low, nrmse_res_G1$ci_low, nrmse_res_B1$ci_low,
                  nrmse_res_N2$ci_low, nrmse_res_G2$ci_low, nrmse_res_B2$ci_low,
                  nrmse_res_N13$ci_low, nrmse_res_G13$ci_low, nrmse_res_B13$ci_low)
cor_ci_low <- c(cor_res_N1$ci_low, cor_res_G1$ci_low, cor_res_B1$ci_low,
                cor_res_N2$ci_low, cor_res_G2$ci_low, cor_res_B2$ci_low,
                cor_res_N13$ci_low, cor_res_G13$ci_low, cor_res_B13$ci_low)
sscore_ci_low <- c(sscore_res_N1$ci_low, sscore_res_G1$ci_low, sscore_res_B1$ci_low,
                   sscore_res_N2$ci_low, sscore_res_G2$ci_low, sscore_res_B2$ci_low,
                   sscore_res_N13$ci_low, sscore_res_G13$ci_low, sscore_res_B13$ci_low)
spcor_ci_low <- c(spcor_res_N1$ci_low, spcor_res_G1$ci_low, spcor_res_B1$ci_low,
                  spcor_res_N2$ci_low, spcor_res_G2$ci_low, spcor_res_B2$ci_low,
                  spcor_res_N13$ci_low, spcor_res_G13$ci_low, spcor_res_B13$ci_low)
acc_ci_low <- c(acc_res_N1$ci_low, acc_res_G1$ci_low, acc_res_B1$ci_low,
                acc_res_N2$ci_low, acc_res_G2$ci_low, acc_res_B2$ci_low,
                acc_res_N13$ci_low, acc_res_G13$ci_low, acc_res_B13$ci_low)
nrmseA_ci_low <- c(nrmseA_res_N1$ci_low, nrmseA_res_G1$ci_low, nrmseA_res_B1$ci_low,
                   nrmseA_res_N2$ci_low, nrmseA_res_G2$ci_low, nrmseA_res_B2$ci_low,
                   nrmseA_res_N13$ci_low, nrmseA_res_G13$ci_low, nrmseA_res_B13$ci_low)

nrmse_ci_high <- c(nrmse_res_N1$ci_high, nrmse_res_G1$ci_high, nrmse_res_B1$ci_high,
                   nrmse_res_N2$ci_high, nrmse_res_G2$ci_high, nrmse_res_B2$ci_high,
                   nrmse_res_N13$ci_high, nrmse_res_G13$ci_high, nrmse_res_B13$ci_high)
cor_ci_high <- c(cor_res_N1$ci_high, cor_res_G1$ci_high, cor_res_B1$ci_high,
                 cor_res_N2$ci_high, cor_res_G2$ci_high, cor_res_B2$ci_high,
                 cor_res_N13$ci_high, cor_res_G13$ci_high, cor_res_B13$ci_high)
sscore_ci_high <- c(sscore_res_N1$ci_high, sscore_res_G1$ci_high, sscore_res_B1$ci_high,
                    sscore_res_N2$ci_high, sscore_res_G2$ci_high, sscore_res_B2$ci_high,
                    sscore_res_N13$ci_high, sscore_res_G13$ci_high, sscore_res_B13$ci_high)
spcor_ci_high <- c(spcor_res_N1$ci_high, spcor_res_G1$ci_high, spcor_res_B1$ci_high,
                   spcor_res_N2$ci_high, spcor_res_G2$ci_high, spcor_res_B2$ci_high,
                   spcor_res_N13$ci_high, spcor_res_G13$ci_high, spcor_res_B13$ci_high)
acc_ci_high <- c(acc_res_N1$ci_high, acc_res_G1$ci_high, acc_res_B1$ci_high,
                 acc_res_N2$ci_high, acc_res_G2$ci_high, acc_res_B2$ci_high,
                 acc_res_N13$ci_high, acc_res_G13$ci_high, acc_res_B13$ci_high)
nrmseA_ci_high <- c(nrmseA_res_N1$ci_high, nrmseA_res_G1$ci_high, nrmseA_res_B1$ci_high,
                    nrmseA_res_N2$ci_high, nrmseA_res_G2$ci_high, nrmseA_res_B2$ci_high,
                    nrmseA_res_N13$ci_high, nrmseA_res_G13$ci_high, nrmseA_res_B13$ci_high)

name <- c("N1", "G1", "B1", "N2", "G2", "B2", "N13", "G13", "B13")

predictors <- c(rep("CAPE", 3), rep("CAPE x Pr", 3), rep("RH + Rsds + T + Ps + Pr + U10", 3))


model <- rep(c("LM", "Gamma GLM", "Gamma Bayesian"),3)

training_period <- rep("2002-2011", 9)

performance_2001_2012 <- data.frame(cbind(name, predictors, model, training_period, nrmse, nrmse_ci_low, nrmse_ci_high, cor, cor_ci_low, cor_ci_high, sscore, sscore_ci_low, sscore_ci_high, spcor, spcor_ci_low, spcor_ci_high, acc, acc_ci_low, acc_ci_high, nrmseA, nrmseA_ci_low, nrmseA_ci_high))

```

Store Intercepts and Coefficints
```{r}
models <- c("N1", "N2", "N13", "G1", "G2", "G13", "B1_alpha", "B1_beta", "B2_alpha", "B2_beta", "B13_alpha", "B13_beta")

coefficients <- data.frame(cbind(models, rbind(c(N1_coefficients, NA, NA, NA, NA, NA), 
                                               c(N2_coefficients, NA, NA, NA, NA, NA), 
                                               N13_coefficients, 
                                               c(G1_coefficients, NA, NA, NA, NA, NA), 
                                               c(G2_coefficients, NA, NA, NA, NA, NA),
                                               G13_coefficients, 
                                               c(B1_alpha_coefficients, NA, NA, NA, NA, NA), 
                                               c(B2_alpha_coefficients, NA, NA, NA, NA, NA),
                                               B13_alpha_coefficients, 
                                               c(B1_beta_coefficients, NA, NA, NA, NA, NA), 
                                               c(B2_beta_coefficients, NA, NA, NA, NA, NA),
                                               B13_beta_coefficients)))

colnames(coefficients) <- c("model", "a", "b", "c", "d", "e", "f", "g")

head(coefficients)

table_file_path <- file.path(table_root, "lightning_model_coefficients_2002_2011.csv")
write.csv(coefficients, table_file_path)
```

#---------------------
#---------------------
2012-2018
#---------------------
#---------------------
```{r}
drivers_file_path <- file.path(drivers_root, "lightning_prediction_drivers_2012_2018.csv")
df <- read.csv(drivers_file_path)[2:13]

colnames(df) <- c("year", "lon", "lat", "strikes", "tair", "precip", "wind", "swr", "sp", "rh", "cape", "cxp")
head(df)
```


2. Split into training and testing data
```{r, echo=FALSE}
## 80% of the sample size
sample_size <- floor(0.8 * nrow(df))

## set the seed to make partition reproducible
set.seed(123)
train_index <- sample(seq_len(nrow(df)), size = sample_size)

train <- df[train_index, ]
test <- df[-train_index, ]

```


#Standardize the data
```{r}
mu_precip <- mean(train$precip)
sigma_precip <- sd(train$precip)
mu_tair <- mean(train$tair)
sigma_tair <- sd(train$tair)
mu_wind <- mean(train$wind)
sigma_wind <- sd(train$wind)
mu_swr <- mean(train$swr)
sigma_swr <- sd(train$swr)
mu_sp <- mean(train$sp)
sigma_sp <- sd(train$sp)
mu_rh <- mean(train$rh)
sigma_rh <- sd(train$rh)
mu_cape <- mean(train$cape)
sigma_cape <- sd(train$cape)
mu_cxp <- mean(train$cxp)
sigma_cxp <- sd(train$cxp)
#-----

train <- train %>%
  mutate(precip = (precip-mu_precip)/sigma_precip) %>%
  mutate(tair = (tair-mu_tair)/sigma_tair) %>%
  mutate(wind = (wind-mu_wind)/sigma_wind) %>%
  mutate(swr = (swr-mu_swr)/sigma_swr) %>%
  mutate(sp = (sp-mu_sp)/sigma_sp) %>%
  mutate(rh = (rh-mu_rh)/sigma_rh) %>%
  mutate(cape = (cape-mu_cape)/sigma_cape) %>%
  mutate(cxp = (cxp-mu_cxp)/sigma_cxp)

train <- train[,c("strikes","precip", "tair", "wind", "swr", "sp", "rh", "cape", "cxp")]
head(train)


test <- test %>%
  mutate(precip = (precip-mu_precip)/sigma_precip) %>%
  mutate(tair = (tair-mu_tair)/sigma_tair) %>%
  mutate(wind = (wind-mu_wind)/sigma_wind) %>%
  mutate(swr = (swr-mu_swr)/sigma_swr) %>%
  mutate(sp = (sp-mu_sp)/sigma_sp) %>%
  mutate(rh = (rh-mu_rh)/sigma_rh) %>%
  mutate(cape = (cape-mu_cape)/sigma_cape) %>%
  mutate(cxp = (cxp-mu_cxp)/sigma_cxp) %>%
  mutate(lon = lon) %>%
  mutate(lat = lat)

test <- test[,c("lon", "lat", "year", "strikes", "precip", "tair", "wind", "swr", "sp", "rh", "cape", "cxp")]

plot(test)

hist(test$strikes)

```


Linear Model
```{r}
#cape
N1 <- lm(strikes ~ cape,
                 data = train)

a <- N1$coefficients[1]
b <- N1$coefficients[2]


N1_coefficients <- c(a, b)

y_pred_N1 <- a + b*test$cape 

nrmse_res_N1  <- metric_nrmse_ci(test$strikes, y_pred_N1)
cor_res_N1   <- metric_cor_ci(test$strikes, y_pred_N1)
sscore_res_N1 <- metric_sscore_ci(test$strikes, y_pred_N1, nbins = 15)
spcor_res_N1  <- metric_spatial_cor_ci(test, y_pred_N1)
acc_res_N1  <- metric_acc_ci(test$year, test$strikes, y_pred_N1)
nrmseA_res_N1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N1)

#CAPE x Precip (cxp)
N2 <- lm(strikes ~ cxp,
                 data = train)

a <- N2$coefficients[1]
b <- N2$coefficients[2]

N2_coefficients <- c(a, b)

y_pred_N2 <- a + b*test$cxp

nrmse_res_N2  <- metric_nrmse_ci(test$strikes, y_pred_N2)
cor_res_N2    <- metric_cor_ci(test$strikes, y_pred_N2)
sscore_res_N2 <- metric_sscore_ci(test$strikes, y_pred_N2, nbins = 15)
spcor_res_N2  <- metric_spatial_cor_ci(test, y_pred_N2)
acc_res_N2  <- metric_acc_ci(test$year, test$strikes, y_pred_N2)
nrmseA_res_N2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N2)

#swr + tair + rh + wind + precip +sp
N13 <- lm(strikes ~ swr + tair + rh + wind + precip +sp,
                 data = train)

a <- N13$coefficients[1]
b <- N13$coefficients[2]
c <- N13$coefficients[3]
d <- N13$coefficients[4]
e <- N13$coefficients[5]
f <- N13$coefficients[6]
g <- N13$coefficients[7]

N13_coefficients <- c(a, b, c, d, e, f, g)

y_pred_N13 <- a + b*test$swr + c*test$tair + d*test$rh +e*test$wind + f*test$precip + g*test$sp

nrmse_res_N13  <- metric_nrmse_ci(test$strikes, y_pred_N13)
cor_res_N13    <- metric_cor_ci(test$strikes, y_pred_N13)
sscore_res_N13 <- metric_sscore_ci(test$strikes, y_pred_N13, nbins = 15)
spcor_res_N13  <- metric_spatial_cor_ci(test, y_pred_N13)
acc_res_N13  <- metric_acc_ci(test$year, test$strikes, y_pred_N13)
nrmseA_res_N13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_N13)

```

Gamma GLM
```{r}

epsilon <- 1e-4  # small value
train$strikes<- train$strikes + epsilon

#cape
G1 <- glm(strikes ~ cape,
                 data = train,
                 family = Gamma(link =log))

a <- G1$coefficients[1]
b <- G1$coefficients[2]

G1_coefficients <- c(a, b)

y_pred_G1 <- exp(a + b*test$cape)

nrmse_res_G1  <- metric_nrmse_ci(test$strikes, y_pred_G1)
cor_res_G1    <- metric_cor_ci(test$strikes, y_pred_G1)
sscore_res_G1 <- metric_sscore_ci(test$strikes, y_pred_G1, nbins = 15)
spcor_res_G1  <- metric_spatial_cor_ci(test, y_pred_G1)
acc_res_G1  <- metric_acc_ci(test$year, test$strikes, y_pred_G1)
nrmseA_res_G1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G1)

#CAPE x precip (cxp)
G2 <- glm(strikes ~ cxp,
                 data = train,
                 family = Gamma(link =log))

a <- G2$coefficients[1]
b <- G2$coefficients[2]

G2_coefficients <- c(a, b)

y_pred_G2 <- exp(a + b*test$cape)

nrmse_res_G2  <- metric_nrmse_ci(test$strikes, y_pred_G2)
cor_res_G2    <- metric_cor_ci(test$strikes, y_pred_G2)
sscore_res_G2 <- metric_sscore_ci(test$strikes, y_pred_G2, nbins = 15)
spcor_res_G2  <- metric_spatial_cor_ci(test, y_pred_G2)
acc_res_G2  <- metric_acc_ci(test$year, test$strikes, y_pred_G2)
nrmseA_res_G2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G2)

#swr + tair + rh + wind + precip +sp
G13 <- glm(strikes ~ swr + tair + rh + wind + precip + sp,
                 data = train,
                 family = Gamma(link =log))

a <- G13$coefficients[1]
b <- G13$coefficients[2]
c <- G13$coefficients[3]
d <- G13$coefficients[4]
e <- G13$coefficients[5]
f <- G13$coefficients[6]
g <- G13$coefficients[7]

G13_coefficients <- c(a, b, c, d, e, f, g)

y_pred_G13 <- exp(a + b*test$swr + c*test$tair + d*test$rh +e*test$wind + f*test$precip + g*test$sp)

nrmse_res_G13  <- metric_nrmse_ci(test$strikes, y_pred_G13)
cor_res_G13    <- metric_cor_ci(test$strikes, y_pred_G13)
sscore_res_G13 <- metric_sscore_ci(test$strikes, y_pred_G13, nbins = 15)
spcor_res_G13  <- metric_spatial_cor_ci(test, y_pred_G13)
acc_res_G13  <- metric_acc_ci(test$year, test$strikes, y_pred_G13)
nrmseA_res_G13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_G13)
```

# B1: CAPE

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: CAPE
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * CAPE,
    beta <- a_beta + b_beta * CAPE,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  CAPE = train$cape
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1,
       a_beta = 1, b_beta = 0.1)
}

if (!file.exists(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2012_2018.rds"))) {
  B1 <- stan(file = "gamma_cape.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B1, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2012_2018.rds"))
    } else {
      B1 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B1_seasonal_2012_2018.rds"))
    }

```


```{r}

#print(B1, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B1)

parameter_stats <- summary(B1)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]

a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]

B1_alpha_coefficients <- c(a_alpha, b_alpha)

B1_beta_coefficients <- c(a_beta, b_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B1)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha

a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$cape)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$cape[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$cape[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  cape = test$cape,            #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$cape, predictions_df$sampled_pred, color=test$sp, main="Lightning Predictions")
plot(test$cape, test$strikes, color=test$sp, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B1 <- predictions_df$sampled_pred
```

```{r}
nrmse_res_B1  <- metric_nrmse_ci(test$strikes, y_pred_B1)
cor_res_B1    <- metric_cor_ci(test$strikes, y_pred_B1)
sscore_res_B1 <- metric_sscore_ci(test$strikes, y_pred_B1, nbins = 15)
spcor_res_B1  <- metric_spatial_cor_ci(test, y_pred_B1)
acc_res_B1  <- metric_acc_ci(test$year, test$strikes, y_pred_B1)
nrmseA_res_B1 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B1)
```

# B2: CAPE x Precipitation (cxp)

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: shortwave radiation,  temperature, relative humidity, wind, precipitation, and surface pressure
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * CXP,
    beta <- a_beta + b_beta * CXP,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  CXP = train$cxp
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1,
       a_beta = 1, b_beta = 0.1)
}

if (!file.exists(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2012_2018.rds"))) {
  B2 <- stan(file = "gamma_cxp.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B2, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2012_2018.rds"))
    } else {
      B2 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B2_seasonal_2012_2018.rds"))
    }

```


```{r}

#print(B2, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B2)

parameter_stats <- summary(B2)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]

a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]

B2_alpha_coefficients <- c(a_alpha, b_alpha)

B2_beta_coefficients <- c(a_beta, b_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B2)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha

a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$cxp)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$cxp[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$cxp[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  cxp = test$cxp,           #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$cxp, predictions_df$sampled_pred, main="Lightning Predictions")
plot(test$cxp, test$strikes, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B2 <- predictions_df$sampled_pred
```

```{r}
nrmse_res_B2  <- metric_nrmse_ci(test$strikes, y_pred_B2)
cor_res_B2    <- metric_cor_ci(test$strikes, y_pred_B2)
sscore_res_B2 <- metric_sscore_ci(test$strikes, y_pred_B2, nbins = 15)
spcor_res_B2  <- metric_spatial_cor_ci(test, y_pred_B2)
acc_res_B2  <- metric_acc_ci(test$year, test$strikes, y_pred_B2)
nrmseA_res_B2 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B2)
```

# B13: Short Wave Radiation + Temperature + Precipitation + Wind + Precipitation + Surface Pressure

Lightning strike rate modeled from a gamma distribution where mu and sigma are dependent on climate: shortwave radiation,  temperature, relative humidity, wind, precipitation, and surface pressure
. 
    S ~ dgamma( alpha , beta ),
    alpha <- a_alpha + b_alpha * SWR + c_alpha * T + d_alpha * RH + e_alpha * W + f_alpha * P + g_alpha * SP,
    beta <- a_beta + b_beta * SWR + c_beta * T + d_beta * RH + e_beta * W + f_beta * P + g_beta * SP,
    
```{r, echo=FALSE, results='hide', include=FALSE}

dat <- list(
  N = nrow(train),
  S = train$strikes,
  SWR = train$swr,
  T = train$tair,
  RH = train$rh,
  W = train$wind,
  P = train$precip,
  SP = train$sp
)


# Compile and fit the Stan model
init_list <- function() {
  list(a_alpha = 1, b_alpha = 0.1, c_alpha = 0.1, d_alpha = 0.1, e_alpha = 0.1, f_alpha = 0.1, g_alpha = 0.1,
       a_beta = 1, b_beta = 0.1, c_beta = 0.1, d_beta = 0.1, e_beta = 0.1, f_beta = 0.1, g_beta = 0.1)
}

if (!file.exists(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2012_2018.rds"))) {
  B13 <- stan(file = "gamma_swr_tair_rh_wind_precip_sp.stan", 
            data = dat,
            iter = 2000, 
            chains = 4, 
            init = init_list,
            cores = min(parallel::detectCores(), 4))
    saveRDS(B13, file = file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2012_2018.rds"))
    } else {
      B13 <- readRDS(file.path(rds_rood, "lightning_prediction_bayes_models/compiled_gamma_B13_seasonal_2012_2018.rds"))
    }

```


```{r}

#print(B13, pars = c("a_alpha", "b_alpha", "c_alpha", "d_alpha", "e_alpha", "f_alpha", "g_alpha", "a_beta", "b_beta", "c_beta", "d_beta", "e_beta", "f_beta", "g_beta"))

traceplot(B13)

parameter_stats <- summary(B13)$summary[, c("mean", "sd")][1:14, 1:2]
print(parameter_stats)

summary_table <- data.frame(
  parameter = rownames(parameter_stats),
  mean = parameter_stats[, "mean"],
  sd = parameter_stats[, "sd"]
)

ggplot(summary_table, aes(x = parameter, y = mean)) +
  geom_point(color = "blue", size = 3) +  # Point for the mean
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.2) +  # Error bars for SD
  labs(y = "Parameter Estimate", x = "Parameter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Parameter Means and SDs from Stan Model")

#store mean posterior sample for coefficient table
a_alpha <- summary_table["a_alpha", "mean"]
b_alpha <- summary_table["b_alpha", "mean"]
c_alpha <- summary_table["c_alpha", "mean"]
d_alpha <- summary_table["d_alpha", "mean"]
e_alpha <- summary_table["e_alpha", "mean"]
f_alpha <- summary_table["f_alpha", "mean"]
g_alpha <- summary_table["g_alpha", "mean"]
a_beta <- summary_table["a_beta", "mean"]
b_beta <- summary_table["b_beta", "mean"]
c_beta <- summary_table["c_beta", "mean"]
d_beta <- summary_table["d_beta", "mean"]
e_beta <- summary_table["e_beta", "mean"]
f_beta <- summary_table["f_beta", "mean"]
g_beta <- summary_table["g_beta", "mean"]

B13_alpha_coefficients <- c(a_alpha, b_alpha, c_alpha, d_alpha, e_alpha, f_alpha, g_alpha)

B13_beta_coefficients <- c(a_beta, b_beta, c_beta, d_beta, e_beta, f_beta, g_beta)
```

Make predictions
```{r}
#Extract posterior samples
posterior_samples <- as.data.frame(B13)

#Extract individual parameters
a_alpha_samples <- posterior_samples$a_alpha
b_alpha_samples <- posterior_samples$b_alpha
c_alpha_samples <- posterior_samples$c_alpha
d_alpha_samples <- posterior_samples$d_alpha
e_alpha_samples <- posterior_samples$e_alpha
f_alpha_samples <- posterior_samples$f_alpha
g_alpha_samples <- posterior_samples$g_alpha
a_beta_samples <- posterior_samples$a_beta
b_beta_samples <- posterior_samples$b_beta
c_beta_samples <- posterior_samples$c_beta
d_beta_samples <- posterior_samples$d_beta
e_beta_samples <- posterior_samples$e_beta
f_beta_samples <- posterior_samples$f_beta
g_beta_samples <- posterior_samples$g_beta

#Number of posterior samples
num_samples <- length(a_alpha_samples)

#Number of test observations
num_test <- length(test$swr)

#Create vectors to store sampled alpha, beta, and predictions
sampled_alpha <- numeric(num_test)
sampled_beta <- numeric(num_test)
sampled_predictions <- numeric(num_test)

#Randomly select a posterior sample for each test observation
posterior_index <- sample(1:num_samples, num_test, replace = TRUE)

for (j in 1:num_test) {
  #Pick a random posterior sample index
  i <- posterior_index[j]  

  #Compute alpha and beta for the test observation
  sampled_alpha[j] <- exp(a_alpha_samples[i] + b_alpha_samples[i] * test$swr[j] + c_alpha_samples[i] * test$tair[j] + d_alpha_samples[i] * test$rh[j] + e_alpha_samples[i] * test$wind[j] + f_alpha_samples[i] * test$precip[j] + g_alpha_samples[i] * test$sp[j])
  sampled_beta[j] <- exp(a_beta_samples[i] + b_beta_samples[i] * test$swr[j] + c_alpha_samples[i] * test$tair[j] + d_alpha_samples[i] * test$rh[j] + e_alpha_samples[i] * test$wind[j] + f_alpha_samples[i] * test$precip[j] + g_alpha_samples[i] * test$sp[j])
  
  #Generate a single prediction from the Gamma distribution
  sampled_predictions[j] <- rgamma(1, shape = sampled_alpha[j], rate = sampled_beta[j])
}

#Create a dataframe with observed and predicted values
predictions_df <- data.frame(
  actual = test$strikes,      #Observed lightning strikes
  swr = test$swr,            #Climate predictor
  tair = test$tair,            #Climate predictor
  rh = test$rh,            #Climate predictor
  wind = test$wind,            #Climate predictor
  precip = test$precip,            #Climate predictor
  sp = test$sp,            #Climate predictor
  alpha = sampled_alpha,      #Sampled alpha for each observation
  beta = sampled_beta,        #Sampled beta for each observation
  sampled_pred = sampled_predictions  # Sampled prediction
)

#add small value back 
predictions_df <- mutate(predictions_df, sampled_pred = sampled_pred + epsilon)

#View predictions
print(head(predictions_df))

hist(predictions_df$sampled_pred, breaks=30, main = "predictions")

hist(test$strikes, breaks=30, main = "observations")

plot(test$swr, predictions_df$sampled_pred, color=test$sp, main="Lightning Predictions")
plot(test$swr, test$strikes, color=test$sp, main="Lightning Observations")
plot(predictions_df$sampled_pred, test$strikes, main="Observed vs. Predicted")
abline(a = 0, b = 1, col = "red", lty = 2, lwd = 2)

#store predictions for figures 
y_pred_B13 <- predictions_df$sampled_pred
```

Performance metrics
```{r}
nrmse_res_B13  <- metric_nrmse_ci(test$strikes, y_pred_B13)
cor_res_B13    <- metric_cor_ci(test$strikes, y_pred_B13)
sscore_res_B13 <- metric_sscore_ci(test$strikes, y_pred_B13, nbins = 15)
spcor_res_B13  <- metric_spatial_cor_ci(test, y_pred_B13)
acc_res_B13  <- metric_acc_ci(test$year, test$strikes, y_pred_B13)
nrmseA_res_B13 <- metric_nrmse_anom_ci(test$year, test$strikes, y_pred_B13)

```


Dataframe of Performance Metrics

```{r}

nrmse <- c(nrmse_res_N1$estimate, nrmse_res_G1$estimate, nrmse_res_B1$estimate, 
           nrmse_res_N2$estimate, nrmse_res_G2$estimate, nrmse_res_B2$estimate, 
           nrmse_res_N13$estimate, nrmse_res_G13$estimate, nrmse_res_B13$estimate)
cor <- c(cor_res_N1$estimate, cor_res_G1$estimate, cor_res_B1$estimate,
         cor_res_N2$estimate, cor_res_G2$estimate, cor_res_B2$estimate,
         cor_res_N13$estimate, cor_res_G13$estimate, cor_res_B13$estimate)
sscore <- c(sscore_res_N1$estimate, sscore_res_G1$estimate, sscore_res_B1$estimate,
            sscore_res_N2$estimate, sscore_res_G2$estimate, sscore_res_B2$estimate,
            sscore_res_N13$estimate, sscore_res_G13$estimate, sscore_res_B13$estimate)
spcor <- c(spcor_res_N1$estimate, spcor_res_G1$estimate, spcor_res_B1$estimate,
           spcor_res_N2$estimate, spcor_res_G2$estimate, spcor_res_B2$estimate,
           spcor_res_N13$estimate, spcor_res_G13$estimate, spcor_res_B13$estimate)
acc <- c(acc_res_N1$estimate, acc_res_G1$estimate, acc_res_B1$estimate, 
         acc_res_N2$estimate, acc_res_G2$estimate, acc_res_B2$estimate,
         acc_res_N13$estimate, acc_res_G13$estimate, acc_res_B13$estimate)
nrmseA <- c(nrmseA_res_N1$estimate, nrmseA_res_G1$estimate, nrmseA_res_B1$estimate,
            nrmseA_res_N2$estimate, nrmseA_res_G2$estimate, nrmseA_res_B2$estimate,
            nrmseA_res_N13$estimate, nrmseA_res_G13$estimate, nrmseA_res_B13$estimate)

nrmse_ci_low <- c(nrmse_res_N1$ci_low, nrmse_res_G1$ci_low, nrmse_res_B1$ci_low,
                  nrmse_res_N2$ci_low, nrmse_res_G2$ci_low, nrmse_res_B2$ci_low,
                  nrmse_res_N13$ci_low, nrmse_res_G13$ci_low, nrmse_res_B13$ci_low)
cor_ci_low <- c(cor_res_N1$ci_low, cor_res_G1$ci_low, cor_res_B1$ci_low,
                cor_res_N2$ci_low, cor_res_G2$ci_low, cor_res_B2$ci_low,
                cor_res_N13$ci_low, cor_res_G13$ci_low, cor_res_B13$ci_low)
sscore_ci_low <- c(sscore_res_N1$ci_low, sscore_res_G1$ci_low, sscore_res_B1$ci_low,
                   sscore_res_N2$ci_low, sscore_res_G2$ci_low, sscore_res_B2$ci_low,
                   sscore_res_N13$ci_low, sscore_res_G13$ci_low, sscore_res_B13$ci_low)
spcor_ci_low <- c(spcor_res_N1$ci_low, spcor_res_G1$ci_low, spcor_res_B1$ci_low,
                  spcor_res_N2$ci_low, spcor_res_G2$ci_low, spcor_res_B2$ci_low,
                  spcor_res_N13$ci_low, spcor_res_G13$ci_low, spcor_res_B13$ci_low)
acc_ci_low <- c(acc_res_N1$ci_low, acc_res_G1$ci_low, acc_res_B1$ci_low,
                acc_res_N2$ci_low, acc_res_G2$ci_low, acc_res_B2$ci_low,
                acc_res_N13$ci_low, acc_res_G13$ci_low, acc_res_B13$ci_low)
nrmseA_ci_low <- c(nrmseA_res_N1$ci_low, nrmseA_res_G1$ci_low, nrmseA_res_B1$ci_low,
                   nrmseA_res_N2$ci_low, nrmseA_res_G2$ci_low, nrmseA_res_B2$ci_low,
                   nrmseA_res_N13$ci_low, nrmseA_res_G13$ci_low, nrmseA_res_B13$ci_low)

nrmse_ci_high <- c(nrmse_res_N1$ci_high, nrmse_res_G1$ci_high, nrmse_res_B1$ci_high,
                   nrmse_res_N2$ci_high, nrmse_res_G2$ci_high, nrmse_res_B2$ci_high,
                   nrmse_res_N13$ci_high, nrmse_res_G13$ci_high, nrmse_res_B13$ci_high)
cor_ci_high <- c(cor_res_N1$ci_high, cor_res_G1$ci_high, cor_res_B1$ci_high,
                 cor_res_N2$ci_high, cor_res_G2$ci_high, cor_res_B2$ci_high,
                 cor_res_N13$ci_high, cor_res_G13$ci_high, cor_res_B13$ci_high)
sscore_ci_high <- c(sscore_res_N1$ci_high, sscore_res_G1$ci_high, sscore_res_B1$ci_high,
                    sscore_res_N2$ci_high, sscore_res_G2$ci_high, sscore_res_B2$ci_high,
                    sscore_res_N13$ci_high, sscore_res_G13$ci_high, sscore_res_B13$ci_high)
spcor_ci_high <- c(spcor_res_N1$ci_high, spcor_res_G1$ci_high, spcor_res_B1$ci_high,
                   spcor_res_N2$ci_high, spcor_res_G2$ci_high, spcor_res_B2$ci_high,
                   spcor_res_N13$ci_high, spcor_res_G13$ci_high, spcor_res_B13$ci_high)
acc_ci_high <- c(acc_res_N1$ci_high, acc_res_G1$ci_high, acc_res_B1$ci_high,
                 acc_res_N2$ci_high, acc_res_G2$ci_high, acc_res_B2$ci_high,
                 acc_res_N13$ci_high, acc_res_G13$ci_high, acc_res_B13$ci_high)
nrmseA_ci_high <- c(nrmseA_res_N1$ci_high, nrmseA_res_G1$ci_high, nrmseA_res_B1$ci_high,
                    nrmseA_res_N2$ci_high, nrmseA_res_G2$ci_high, nrmseA_res_B2$ci_high,
                    nrmseA_res_N13$ci_high, nrmseA_res_G13$ci_high, nrmseA_res_B13$ci_high)

name <- c("N1", "G1", "B1", "N2", "G2", "B2", "N13", "G13", "B13")

predictors <- c(rep("CAPE", 3), rep("CAPE x Pr", 3), rep("RH + Rsds + T + Ps + Pr + U10", 3))


model <- rep(c("LM", "Gamma GLM", "Gamma Bayesian"),3)

training_period <- rep("2012-2018", 9)

performance_2012_2018 <- data.frame(cbind(name, predictors, model, training_period, nrmse, nrmse_ci_low, nrmse_ci_high, cor, cor_ci_low, cor_ci_high, sscore, sscore_ci_low, sscore_ci_high, spcor, spcor_ci_low, spcor_ci_high, acc, acc_ci_low, acc_ci_high, nrmseA, nrmseA_ci_low, nrmseA_ci_high))


performance <- rbind(performance_2001_2012, performance_2012_2018)

table_file_path <- file.path(table_root, "lightning_model_performance.csv")
write.csv(performance, table_file_path)
```

Store Intercepts and Coefficints
```{r}
models <- c("N1", "N2", "N13", "G1", "G2", "G13", "B1_alpha", "B1_beta", "B2_alpha", "B2_beta", "B13_alpha", "B13_beta")

coefficients <- data.frame(cbind(models, rbind(c(N1_coefficients, NA, NA, NA, NA, NA), 
                                               c(N2_coefficients, NA, NA, NA, NA, NA), 
                                               N13_coefficients, 
                                               c(G1_coefficients, NA, NA, NA, NA, NA), 
                                               c(G2_coefficients, NA, NA, NA, NA, NA),
                                               G13_coefficients, 
                                               c(B1_alpha_coefficients, NA, NA, NA, NA, NA), 
                                               c(B2_alpha_coefficients, NA, NA, NA, NA, NA),
                                               B13_alpha_coefficients, 
                                               c(B1_beta_coefficients, NA, NA, NA, NA, NA), 
                                               c(B2_beta_coefficients, NA, NA, NA, NA, NA),
                                               B13_beta_coefficients)))

colnames(coefficients) <- c("model", "a", "b", "c", "d", "e", "f", "g")

head(coefficients)

table_file_path <- file.path(table_root, "lightning_model_coefficients_2012_2018.csv")
write.csv(coefficients, table_file_path)
```