---
title: "Figure 2"
author: "Charlotte Uden"
date: "2026-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(scales)
library(stringr)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths 
```

```{r}
df <- read.csv(file.path(table_root, "lightning_model_performance.csv"))

# Define the desired order
desired_order <- c("CAPE", "CAPE x Pr", "RH + Rsds + T + Ps + Pr + U10")

df <- df %>%
  dplyr::mutate(
    predictors    = factor(predictors, levels = desired_order),
    model      = factor(model, levels = c("LM", "Gamma GLM", "Gamma Bayesian")),
    training_period = factor(training_period, levels = c("2001-2012", "2012-2018"))
  )
```

Note: some spcor esimates fall outside the confidence interval - this is normal. Explanation in a reviewer asks:

Bootstrap confidence intervals were computed from the distribution of statistics over 1,000 resampled test sets. As a result, the original test-set estimate is not constrained to lie within the percentile-based confidence interval, particularly when the sampling distribution is skewed. We interpret these intervals as approximations of the sampling uncertainty, rather than as strict bounds on the observed test statistic.

```{r}
# A base theme for a clean, publication-style look
base_theme <- theme_bw(base_size = 11) +
  theme(
    #panel.grid.major.y = element_blank(),
    #panel.grid.minor   = element_blank(),
    axis.title.y       = element_blank(),
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.title       = element_text(size = 10),
    legend.text        = element_text(size = 9)
  )

# ------------------ Normalized RMSE ------------------ #
p_nrmse <- ggplot(df,
                aes(x = nrmse, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = nrmse_ci_low,
                     xmax = nrmse_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_reverse(
    breaks = seq(0, 14, by = 1),
    labels = label_number(accuracy = 1)) +
  coord_cartesian(xlim = c(14, 0)) + 
  scale_y_discrete(labels = function(x) str_wrap(x, width = 15)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "nRMSE",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(a)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =10, b = 10, l = 30)) +
  theme(legend.position = "none")

# ------------------ Correlation ------------------ #
p_cor <- ggplot(df,
                aes(x = cor, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = cor_ci_low,
                     xmax = cor_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_continuous(
    limits = c(-0.1, 0.7),
    breaks = seq(-0.1, 0.7, by = 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "Correlation",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(b)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =30, b = 10, l = 10)) +
  scale_y_discrete(position = "right",
                   labels = function(x) str_wrap(x, width = 15)) +
  theme(legend.position = "none")

# ------------------ S-SCORE------------------ #
p_sscore <- ggplot(df,
                  aes(x = sscore, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = sscore_ci_low,
                     xmax = sscore_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_continuous(
    limits = c(0.5, 1),
    breaks = seq(0.5, 1, by = 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 15)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "S-score",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(c)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =10, b = 10, l = 30)) +
  theme(legend.position = "none")

# ------------------ SPATIAL CORRELATION ------------------ #
p_spcor <- ggplot(df,
                  aes(x = spcor, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = spcor_ci_low,
                     xmax = spcor_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_continuous(
    breaks = seq(-0.1, 0.7, by = 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "Spatial Correlation",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(d)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =30, b = 10, l = 10)) +
  scale_y_discrete(position = "right",
                   labels = function(x) str_wrap(x, width = 15)) +
  theme(legend.position = "none")

# ------------------ ANOMALY CORRELATION COEFFICIENT ------------------ #
p_acc <- ggplot(df,
                  aes(x = acc, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = acc_ci_low,
                     xmax = acc_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_continuous(
    breaks = seq(-1, 1, by = 0.1),
    labels = label_number(accuracy = 0.1)
  ) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 15)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "ACC",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(e)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =10, b = 10, l = 30)) +
  theme(legend.position = "none")

# ------------------ nRMSE of ANOMALIES ------------------ #
p_nrmseA <- ggplot(df,
                  aes(x = nrmseA, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = nrmseA_ci_low,
                     xmax = nrmseA_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_reverse(
    breaks = seq(0, 5, by = 0.5),
    labels = label_number(accuracy = 0.1)
  ) +
  coord_cartesian(xlim = c(5, 0)) + 
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "nRMSE of Anomalies",
    y = NULL,
    color = "Approach",
    shape = "Training period"
  ) +
  ggtitle("(f)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin=ggplot2::margin(t = 10, r =30, b = 10, l = 10)) +
  scale_y_discrete(position = "right",
                   labels = function(x) str_wrap(x, width = 15)) +
  theme(legend.position = "none")


# ------------------ COMBINE PANELS ------------------ #

# Extract the legend
#----
# Function to extract the legend from a ggplot object
extract_legend <- function(my_plot) {
  g <- ggplotGrob(my_plot)  # Convert plot to grob
  legend <- g$grobs[which(sapply(g$grobs, function(x) x$name) == "guide-box")][[1]]
  return(legend)  # Returns the legend as a grob
}

legend_plot <- ggplot(df,
                  aes(x = nrmseA, y = predictors,
                    color = model,
                    shape = training_period)) +
  geom_point(size = 2.8, alpha = 0.9,
                 position = position_dodge(width = 0.6)) +
  # --- confidence intervals ---
  geom_errorbarh(aes(xmin = nrmseA_ci_low,
                     xmax = nrmseA_ci_high),
                 height = 0.25,
                 alpha = 0.6,
                 linewidth = 0.5,
                 position = position_dodge(width = 0.6)) +
  scale_x_continuous(
    breaks = seq(0, 5, by = 0.5),
    labels = label_number(accuracy = 0.1)
  ) +
  coord_cartesian(xlim = c(-2, 14)) + 
  scale_color_brewer(palette = "Dark2") +
  labs(
    x = "nRMSE of Anomalies",
    y = NULL,
    color = "Approach",
    shape = "Years"
  ) +
  ggtitle("(f)") +
  base_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_y_discrete(position = "right",
                   labels = function(x) str_wrap(x, width = 15)) +
  theme(
    legend.box = "horizontal",
    legend.text = element_text(size = 10),
    legend.key.height = unit(0.5, "cm"), # Add a small height back
    legend.key.width = unit(1.5, "cm"),  # Slightly reduced width
    legend.spacing.x = unit(0.5, "cm")
  ) +
  guides(
    color = guide_legend(ncol = 3),
    shape = guide_legend(ncol = 3) 
  )

#----

legend_grob <- extract_legend(legend_plot)


p <- grid.arrange(p_nrmse, p_cor, p_sscore, p_spcor, p_acc, p_nrmseA, legend_grob, nrow=4, ncol=2, heights=c(1,1,1,0.2), widths=c(1, 1), layout_matrix=rbind(c(1, 2), c(3, 4), c(5, 6), c(7,7)))
```

```{r}
ggsave(file.path(figure_root, "Figure_2.png"), plot = p, width = 10, height = 9, dpi = 1100, units = "in")
```

