---
title: "Figure_6"
author: "Charlotte Uden"
date: "2026-01-30"
output: html_document
---

Figure 6. Interannual variability of observed fire frequency and simulated fire probability in Alaska, 2002-2011. Yellow circles show annual mean observed fire rate (number of fire days/total number of days) with error bars representing 95% binomial confidence intervals around the annual mean fire occurrence. Blue triangles show annual mean predicted fire probability with error bars representing bootstrap 95% confidence interval of the annual mean, via resampling cells with replacement. The dotted horizontal line marks the ten-year observed mean lightning-caused fire rate (0.00067) and provides a reference for interannual anomalies. 


```{r}
library(tidyr)
library(patchwork)
library(dplyr)
library(ggplot2)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths 
```

```{r}
df <- read.csv(file.path(predictions_root, "drivers_predictions_and_observations_2001-2012_10x.csv"))
```

#Bars for observed fire rate:

Use a binomial standard error / confidence interval instead of standard deviation. For each year, compute:
mu = mean(fire)
SE = sqrt(mu*(1-mu)/n)
error bar = mu +- 1.96*SE
This gives uncertainty on the mean rate, not spread among daily 0/1 values. 

```{r}

summary_df <- df %>%
  group_by(year) %>%
  summarise(
    mean_fire   = mean(fire, na.rm = TRUE),
    n           = sum(is.finite(fire)),
    .groups = "drop"
  ) %>%
  mutate(
    se_fire = sqrt(mean_fire * (1 - mean_fire) / n),
    lo_fire = pmax(0, mean_fire - 1.96 * se_fire),
    hi_fire = pmin(1, mean_fire + 1.96 * se_fire),
  )



```

#Bars for predicted fire probability:

Bootstrap uncertainty of the annual mean predicted probability of fire. For each year, mean_pred = mean (P_fire) over all gridcell-days
lo, high = bootsrap 95% CI of the annual mean, via resampling cells with replacement

```{r}
bootstrap_annual_mean_pred <- function(df, p_col = "P_fire", year_col = "year",
                                       cluster_col = "cell_id",
                                       B = 1000, conf = 0.95, seed = 1) {
  stopifnot(all(c(p_col, year_col, cluster_col) %in% names(df)))

  # Keep only needed columns, drop NAs
  d <- df[, c(year_col, cluster_col, p_col)]
  d <- d[is.finite(d[[p_col]]) & is.finite(d[[year_col]]), , drop = FALSE]

  # Use data.table for speed/memory
  if (!requireNamespace("data.table", quietly = TRUE)) {
    stop("Please install data.table for this function: install.packages('data.table')")
  }
  dt <- data.table::as.data.table(d)
  data.table::setnames(dt, c(year_col, cluster_col, p_col), c("year", "cluster", "p"))

  # Precompute row indices for each (year, cluster)
  idx_by_cluster <- dt[, .(idx = list(.I)), by = .(year, cluster)]

  # For each year, store: unique clusters + list of indices per cluster
  years <- sort(unique(idx_by_cluster$year))
  alpha <- (1 - conf) / 2

  set.seed(seed)

  out <- lapply(years, function(yy) {
    sub <- idx_by_cluster[year == yy]
    clusters <- sub$cluster
    idx_list <- sub$idx
    K <- length(clusters)

    # Point estimate (no resampling)
    mean_pred <- mean(dt[year == yy, p])

    # Bootstrap
    boot_means <- numeric(B)
    for (b in seq_len(B)) {
      # resample clusters with replacement
      samp <- sample.int(K, size = K, replace = TRUE)

      # concatenate row indices for sampled clusters
      idx <- unlist(idx_list[samp], use.names = FALSE)

      boot_means[b] <- mean(dt$p[idx])
    }

    qs <- stats::quantile(boot_means, probs = c(alpha, 1 - alpha), names = FALSE, type = 7)

    data.frame(
      year = yy,
      mean_pred = mean_pred,
      lo_pred = qs[1],
      hi_pred = qs[2]
    )
  })

  out <- do.call(rbind, out)
  out
}

pred_ci <- bootstrap_annual_mean_pred(df, p_col = "i.P_fire", cluster_col = "cell_id",
                                      B = 1000, seed = 123)

head(pred_ci)
```

#merge and convert to long format
```{r}
summary_df <- merge(summary_df, pred_ci, by="year")

summary_long <- summary_df %>%
  transmute(
    year,
    series = "Observed fire rate",
    mean   = mean_fire,
    lo     = lo_fire,
    hi     = hi_fire
  ) %>%
  bind_rows(
    summary_df %>%
      transmute(
        year,
        series = "Predicted fire probability",
        mean   = mean_pred,
        lo     = lo_pred,
        hi     = hi_pred
      )
  )
```


```{r}
pd <- position_dodge(width = 0.4)

custom_grid_lines_x <- c(2002.5, 2003.5, 2004.5, 2005.5, 2006.5, 2007.5, 2008.5, 2009.5, 2010.5, 2011.5) 

p <- ggplot(summary_long,
            aes(x = year, y = mean, color = series, shape = series)) +
  
  geom_point(size = 3, position = pd) +
  
  geom_errorbar(aes(ymin = lo, ymax = hi),
                width = 0.2, position = pd) +
  
  scale_color_manual(values = c("Observed fire rate" = "#FFC20A", "Predicted fire probability" = "#0C7BDC"),
                     labels = c("Observed fire rate", "Predicted fire probability")) +
  
  geom_hline(yintercept = sum(df$fire)/nrow(df), linetype = "dotted", color = "black", linewidth = 0.5) +
  
  scale_x_continuous(breaks = summary_df$year) +
  
  theme_bw() +
   geom_vline(xintercept = custom_grid_lines_x, color = "gray", linewidth = 0.1) +

  theme(
    panel.grid.major.x = element_blank(), # Remove default major vertical gridlines
    panel.grid.minor.x = element_blank(), # Remove default minor vertical gridlines
    panel.grid.major.y = element_line(color = "gray", linewidth = 0.1), # Keep/style horizontal gridlines
    panel.grid.minor.y = element_blank()  # Remove minor horizontal gridlines if desired
  ) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  labs(x = "Year", y = "Fire probability / rate") +
  
  scale_y_continuous(labels = scales::label_number(accuracy = 0.0001))

p
```


```{r}
ggsave(file.path(figure_root, "Figure_6.png"), plot = p, width = 8, height = 3, dpi = 1100, units = "in")
```



