---
title: "Figure_C1-2"
author: "Charlotte Uden"
date: "2026-03-13"
output: html_document
---

```{r}
library(randomForest)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidytext)
library(patchwork)
```

#########################
Lightning strike rate
#########################

Figure C1. Ransom forest results testing variable importance in predicting lightning strike rate. Metrics include (a) percent increase in mean square error (%IncMSE) and (b) increase in node impurity (IncNodePurity). Predictor variables include monthly mean wind speed (m/s), relative humidity (rh; %), shortwave radiation (swr; W/m2)), precipitation (precip; mm), surface pressure (sp; Pascal), and temperature (tair; °C). Color indicates the training period. 

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths
```

```{r}
rf_r_strike_1 <- read.csv(file.path(table_root, "lightning_random_forest_results_2002_2011.csv"))
rf_r_strike_1$years <- rep("2002-2011", nrow(rf_r_strike_1))
```


```{r}
rf_r_strike_2 <- read.csv(file.path(table_root, "lightning_random_forest_results_2012_2018.csv"))
rf_r_strike_2$years <- rep("2012-2018", nrow(rf_r_strike_2))

rf_r_strike <- rbind(rf_r_strike_2, rf_r_strike_1)
head(rf_r_strike)
```


```{r}
#-------------------------
# IncMSE plot
#-------------------------
rf_mse <- rf_r_strike %>%
  select(Variable, IncMSE, years) %>%
  group_by(Variable) %>%
  mutate(order_val = mean(IncMSE, na.rm = TRUE)) %>%
  ungroup()

p_mse <- ggplot(rf_mse,
                aes(x = IncMSE,
                    y = reorder(Variable, order_val),
                    fill = years)) +
  geom_hline(
    yintercept = seq_along(unique(rf_mse$Variable)),
    linetype = "dashed",
    color = "grey75",
    linewidth = 0.4
  ) +
  geom_point(
    shape = 21,    
    size = 3,
    color = "black",
    stroke = 0.7
  ) +
  labs(
    title = "(a)",
    x = "%IncMSE",
    y = NULL,
    fill = "Training years"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    strip.background = element_blank(),
    legend.position = "bottom"
  ) + 
  theme(legend.position = "none")

#-------------------------
# IncNodePurity plot
#-------------------------
rf_node <- rf_r_strike %>%
  select(Variable, IncNodePurity, years) %>%
  group_by(Variable) %>%
  mutate(order_val = mean(IncNodePurity, na.rm = TRUE)) %>%
  ungroup()

p_node <- ggplot(rf_node,
                 aes(x = IncNodePurity,
                     y = reorder(Variable, order_val),
                     fill = years)) +
  geom_hline(
    yintercept = seq_along(unique(rf_node$Variable)),
    linetype = "dashed",
    color = "grey75",
    linewidth = 0.4
  ) +
  geom_point(
    shape = 21,         # circle with border
    size = 3,
    color = "black",
    stroke = 0.7
  ) +
  labs(
    title = "(b)",
    x = "IncNodePurity",
    y = NULL,
    fill = "Training years"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    strip.background = element_blank(),
    legend.position = "bottom"
  )

(p_mse + p_node) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r}
ggsave(file.path(figure_root, "Figure_C1.png"), plot = last_plot(), width = 8, height = 4, dpi = 750)
```

#########################
Ignition Efficiency
#########################

Figure C2. Random forest results testing variable importance in predicting lightning ignition efficiency. The x-axis of each plot shows (a) percent increase in mean square error (%IncMSE) and (b) increase in node impurity (IncNodePurity). X-axes display predictor variables: monthly mean wind speed (m/s), relative humidity (rh; %), precipitation (precip; mm), temperature (tair; °C), fine fuel moisture code (FFMC), duff moisture code (DMC), and drought code (DC), % conifer cover (C), % broadleaf cover (B), and % unvegetated/openland (U). Color indicates training period. 

```{r}
rf_p_ignite_1 <- read.csv(file.path(table_root, "ignition_efficiency_random_forest_results_2012_2018.csv"))
rf_p_ignite_1$years <- rep("2012-2018", nrow(rf_p_ignite_1))
```


```{r}
rf_p_ignite_2 <- read.csv(file.path(table_root, "ignition_efficiency_random_forest_results_2002_2011.csv"))
rf_p_ignite_2$years <- rep("2002-2011", nrow(rf_p_ignite_2))

rf_p_ignite <- rbind(rf_p_ignite_2, rf_p_ignite_1)
```


```{r}
#-------------------------
# IncMSE plot
#-------------------------
rf_mse <- rf_p_ignite %>%
  select(Variable, IncMSE, years) %>%
  group_by(Variable) %>%
  mutate(order_val = mean(IncMSE, na.rm = TRUE)) %>%
  ungroup()

p_mse <- ggplot(rf_mse,
                aes(x = IncMSE,
                    y = reorder(Variable, order_val),
                    fill = years)) +
  geom_hline(
    yintercept = seq_along(unique(rf_mse$Variable)),
    linetype = "dashed",
    color = "grey75",
    linewidth = 0.4
  ) +
  geom_point(
    shape = 21,    
    size = 3,
    color = "black",
    stroke = 0.7
  ) +
  labs(
    title = "(a)",
    x = "%IncMSE",
    y = NULL,
    fill = "Training years"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    strip.background = element_blank(),
    legend.position = "bottom"
  ) + 
  theme(legend.position = "none")

#-------------------------
# IncNodePurity plot
#-------------------------
rf_node <- rf_p_ignite %>%
  select(Variable, IncNodePurity, years) %>%
  group_by(Variable) %>%
  mutate(order_val = mean(IncNodePurity, na.rm = TRUE)) %>%
  ungroup()

p_node <- ggplot(rf_node,
                 aes(x = IncNodePurity,
                     y = reorder(Variable, order_val),
                     fill = years)) +
  geom_hline(
    yintercept = seq_along(unique(rf_node$Variable)),
    linetype = "dashed",
    color = "grey75",
    linewidth = 0.4
  ) +
  geom_point(
    shape = 21,         # circle with border
    size = 3,
    color = "black",
    stroke = 0.7
  ) +
  labs(
    title = "(b)",
    x = "IncNodePurity",
    y = NULL,
    fill = "Training years"
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    strip.background = element_blank(),
    legend.position = "bottom"
  )

(p_mse + p_node) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r}
ggsave(file.path(figure_root, "Figure_C2.png"), plot = last_plot(), width = 8, height = 5, dpi = 750)
```



