---
54title: "Figure_8"
author: "Charlotte Uden"
date: "2026-03-06"
output: html_document
---

Figure 8. Climate controls on modeled lightning strike rate, ignition efficiency, and lightning-caused fire probability in Alaska, 2002-2011. Heatmaps show the mean modeled response to combinations of daily near surface climate variables: temperature and relative humidity (a-c), temperature and precipitation (d-f), temperature and wind speed (g-i), relative humidity and precipitation (j-l), precipitation and wind speed (m-o), and relative humidity and wind speed (p-r). These climate variables were selected because they are applied in both the lightning and ignition efficiency models, allowing consistent comparison of their influence across models. Columns correspond to the three modeled components: lightning strike rate (strikes/month/km2; left), ignition efficiency (center), and fire probability (right). Colors indicate the mean model prediction within quantile-based climate bins with warmer colors indicating higher predicted values.


```{r}

library(tidyr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(terra)
library(lubridate)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths 
```

#Read in observations of fire occurance and predictions of r_strike, P_ignite, and P_fire
```{r}
df <- read.csv(file.path(drivers_root, "drivers_predictions_and_observations_2002-2011_10x.csv"))
str(df)
```


Function to preprocess data for plotting
```{r}
plot_preprocess <- function(data, x_var, y_var, target_name, x_bins, y_bins, min_count) {
  
  # --- Clean + transform ---
# Select only the necessary columns for this plot and transform
  plot_df <- data %>%
    dplyr::select(all_of(c(x_var, y_var, target_name))) %>%
    mutate(across(everything(), as.numeric)) %>%
    drop_na()

# Add the log10 target column (floor to avoid -Inf for very small probabilities)
  plot_df <- plot_df %>%
    mutate(
      !!log_target_col := log10(pmax(.data[[target_name]], 1e-15))
      )

# --- Binning and Aggregation for the single heatmap ---

# Calculate breaks for x_var (ie. temperature)
  x_breaks <- quantile(plot_df[[x_var]], probs = seq(0, 1, length.out = x_bins + 1), na.rm = TRUE)
  x_breaks <- unique(x_breaks) # Ensure breaks are unique
  if (length(x_breaks) < 2) { # Fallback if not enough unique breaks
    x_breaks <- pretty(plot_df[[x_var]], n = x_bins)
    }

# Calculate breaks for y_var (ie. precipitation)
  y_breaks <- quantile(plot_df[[y_var]], probs = seq(0, 1, length.out = y_bins + 1), na.rm = TRUE)
  y_breaks <- unique(y_breaks) # Ensure breaks are unique
  if (length(y_breaks) < 2) { # Fallback if not enough unique breaks
    y_breaks <- pretty(plot_df[[y_var]], n = y_bins)
    }

# Create bins and aggregate data
  surf_df <- plot_df %>%
    mutate(
      x_bin = cut(.data[[x_var]], breaks = x_breaks, include.lowest = TRUE, labels = FALSE),
      y_bin = cut(.data[[y_var]], breaks = y_breaks, include.lowest = TRUE, labels = FALSE)
      ) %>%
    group_by(x_bin, y_bin) %>%
    summarise(
      mean_target = mean(.data[[target_name]], na.rm = TRUE),
      mean_log_target = mean(.data[[log_target_col]], na.rm = TRUE),#get mean of log column too
      count = n(),
      .groups = "drop"
      ) %>%
    filter(count >= min_count) %>% # Mask bins with too few samples
    #mutate(count = ifelse(count < min_count, NA, count)) %>% #or make counts less than min_count NA
    #filter(mean_target >= 0.00001) %>%
    # Calculate bin midpoints for plotting
    mutate(
      xmid = (x_breaks[x_bin] + x_breaks[x_bin + 1]) / 2,
      ymid = (y_breaks[y_bin] + y_breaks[y_bin + 1]) / 2
      ) %>%
    mutate(
      xmin = x_breaks[x_bin],
      xmax = x_breaks[x_bin + 1],
      ymin = y_breaks[y_bin],
      ymax = y_breaks[y_bin + 1]
    )
  return(surf_df)
}
```

##########################
Fire Probability
##########################

temp x RH
```{r}
plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="relative_humidity",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_t_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013),
                       guide = "none") +
  labs(
    title = "(c)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Relative humidity (%)",
    fill = "Fire probability"
  )
  
p_fire_t_rh


#temp x precip

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="precipitation",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_t_p <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013),
                       guide = "none") +
  labs(
    title = "(f)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Precipitation (mm)",
    fill = "Fire probability"
  ) + 
  theme_minimal(base_size = 12) 
  
p_fire_t_p


#temp x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="wind",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_t_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013),
                       guide = "none") +
  labs(
    title = "(i)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Wind (m/s)",
    fill = "Fire probability"
  )+theme_minimal(base_size = 12) 
  
p_fire_t_w


#precip x rh

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="precipitation",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_p_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013),
                       guide = "none") +
  labs(
    title = "(l)",
    x = "Relative humidity (%)",
    y = "Precipitation (mm)",
    fill = "Fire probability"
  ) + 
  theme_minimal(base_size = 12) 
  
p_fire_p_rh


#precip x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="wind", 
                            y_var="precipitation",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_p_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013),
                       guide = "none") +
  labs(
    title = "(o)",
    x = "Wind (m/s)",
    y = "Precipitation (mm)",
    fill = "Fire probability"
  ) + 
  theme_minimal(base_size = 12) 
  
p_fire_p_w


#relative humidity x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="wind",
                            target_name="P_fire",
                            x_bins=40, y_bins=40, min_count=200)

p_fire_rh_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.0013)) +
  labs(
    title = "(r)",
    x = "Relative humidity (%)",
    y = "Wind (m/s)",
    fill = "Fire probability      "
  ) + 
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 16)) +
  guides(fill = guide_colorbar(
    title.position = "top",      # Place the title above the color bar
      barwidth = 20,               # Adjust bar width for better horizontal layout
      barheight = 3,
    label.theme = element_text(angle = 45, hjust = 1)
  ))
  
p_fire_rh_w
```



```{r}
p_fire_t_rh |p_fire_t_p | p_fire_t_w | p_fire_p_rh | p_fire_p_w | p_fire_rh_w


```


##########################
Ignition Efficiency
##########################

temp x RH
```{r}
plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="relative_humidity",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_t_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.51),
                       guide = "none"
                       ) +
  labs(
    title = "(b)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Relative humidity (%)",
    fill = "Ignition efficiency"
  )
  
p_ignite_t_rh


#temp x precip

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="precipitation",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_t_p <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.51),
                       guide = "none"
                       ) +
  labs(
    title = "(e)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Precipitation (mm)",
    fill = "Ignition efficiency"
  ) + 
  theme_minimal(base_size = 12) 
  
p_ignite_t_p


#temp x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="wind",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_t_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.51),
                       guide = "none"
                       ) +
  labs(
    title = "(h)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Wind (m/s)",
    fill = "Ignition efficiency"
  )+theme_minimal(base_size = 12) 
  
p_ignite_t_w


#precip x rh

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="precipitation",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_p_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.51),
                       guide = "none"
                       ) +
  labs(
    title = "(k)",
    x = "Relative humidity (%)",
    y = "Precipitation (mm)",
    fill = "Ignition efficiency"
  ) + 
  theme_minimal(base_size = 12) 
  
p_ignite_p_rh


#precip x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="wind", 
                            y_var="precipitation",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_p_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.51),
                       guide = "none"
                       ) +
  labs(
    title = "(n)",
    x = "Wind (m/s)",
    y = "Precipitation (mm)",
    fill = "Ignition efficiency"
  ) + 
  theme_minimal(base_size = 12) 
  
p_ignite_p_w


#relative humidity x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="wind",
                            target_name="P_ignite_prob",
                            x_bins=40, y_bins=40, min_count=200)

p_ignite_rh_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.1),
                       limits = c(0, 0.51)
                       ) +
  labs(
    title = "(q)",
    x = "Relative humidity (%)",
    y = "Wind (m/s)",
    fill = "Ignition efficiency     "
  ) + 
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 16)) +
  guides(fill = guide_colorbar(
    title.position = "top",      # Place the title above the color bar
      barwidth = 20,               # Adjust bar width for better horizontal layout
      barheight = 3,
    label.theme = element_text(angle = 45, hjust = 1)
  ))
  
p_ignite_rh_w
```


```{r}
p_ignite_t_rh |p_ignite_t_p | p_ignite_t_w | p_ignite_p_rh | p_ignite_p_w | p_ignite_rh_w

```


##########################
Strike rate
##########################

temp x RH
```{r}
plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="relative_humidity",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_t_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.06),
                       guide = "none"
                       ) +
  labs(
    title = "(a)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Relative humidity (%)",
    fill = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2))
  )
  
p_r_strike_t_rh


#temp x precip

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="precipitation",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_t_p <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.06),
                       guide = "none"
                       ) +
  labs(
    title = "(d)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Precipitation (mm)",
    fill = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2))
  ) + 
  theme_minimal(base_size = 12) 
  
p_r_strike_t_p


#temp x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="wind",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_t_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.06),
                       guide = "none"
                       ) +
  labs(
    title = "(g)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Wind (m/s)",
    fill = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2))
  ) + theme_minimal(base_size = 12) 
  
p_r_strike_t_w


#precip x rh

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="precipitation",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_p_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.06),
                       guide = "none"
                       ) +
  labs(
    title = "(j)",
    x = "Relative humidity (%)",
    y = "Precipitation (mm)",
    fill = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2))
  ) + 
  theme_minimal(base_size = 12) 
  
p_r_strike_p_rh


#precip x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="wind", 
                            y_var="precipitation",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_p_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.06),
                       guide = "none"
                       ) +
  labs(
    title = "(m)",
    x = "Wind (m/s)",
    y = "Precipitation (mm)",
    fill = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2))
  ) + 
  theme_minimal(base_size = 12) 
  
p_r_strike_p_w


#relative humidity x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="wind",
                            target_name="r_strike",
                            x_bins=40, y_bins=40, min_count=200)

p_r_strike_rh_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.01),
                       limits = c(0, 0.06)
                       ) +
  labs(
    title = "(p)",
    x = "Relative humidity (%)",
    y = "Wind (m/s)",
    fill = "Strike rate     "
  ) + 
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 16)) +
  guides(fill = guide_colorbar(
    title.position = "top",      # Place the title above the color bar
      barwidth = 20,               # Adjust bar width for better horizontal layout
      barheight = 3,
    label.theme = element_text(angle = 45, hjust = 1)
  ))
  
p_r_strike_rh_w
```

```{r}
p_r_strike_t_rh |p_r_strike_t_p | p_r_strike_t_w | p_r_strike_p_rh | p_r_strike_p_w | p_r_strike_rh_w

```

##########################
Observed Fire Rate
##########################

temp x RH
```{r}
plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="relative_humidity",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_t_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005),
                       guide = "none"
                       ) +
  labs(
    title = "(a)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Relative humidity (%)",
    fill = "Observed fire rate"
  )
  
p_obs_fire_t_rh


#temp x precip

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="precipitation",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_t_p <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005),
                       guide = "none"
                       ) +
  labs(
    title = "(b)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Precipitation (mm)",
    fill = "Observed fire rate"
  ) + 
  theme_minimal(base_size = 12) 
  
p_obs_fire_t_p


#temp x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="temperature", 
                            y_var="wind",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_t_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005),
                       guide = "none"
                       ) +
  labs(
    title = "(c)",
    x = expression("Temperature ("*degree*C*")"),
    y = "Wind (m/s)",
    fill = "Observed fire rate"
  )+theme_minimal(base_size = 12) 
  
p_obs_fire_t_w


#precip x rh

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="precipitation",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_p_rh <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005),
                       guide = "none"
                       ) +
  labs(
    title = "(d)",
    x = "Relative humidity (%)",
    y = "Precipitation (mm)",
    fill = "Observed fire rate"
  ) + 
  theme_minimal(base_size = 12) 
  
p_obs_fire_p_rh


#precip x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="wind", 
                            y_var="precipitation",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_p_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005),
                       guide = "none"
                       ) +
  labs(
    title = "(e)",
    x = "Wind (m/s)",
    y = "Precipitation (mm)",
    fill = "Observed fire rate"
  ) + 
  theme_minimal(base_size = 12) 
  
p_obs_fire_p_w


#relative humidity x wind

plot_dat <- plot_preprocess(data=df, 
                            x_var="relative_humidity", 
                            y_var="wind",
                            target_name="fire",
                            x_bins=40, y_bins=40, min_count=200)

p_obs_fire_rh_w <- ggplot(plot_dat, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = mean_target)) +
  geom_rect(color = NA) +
  scale_fill_viridis_c(na.value = "grey50", labels = label_number(accuracy = 0.00001),
                       limits = c(0, 0.005)
                       ) +
  labs(
    title = "(f)",
    x = "Relative humidity (%)",
    y = "Wind (m/s)",
    fill = "Observed fire rate      "
  ) + 
  theme_minimal(base_size = 12) + 
  theme(legend.position = "bottom",
        legend.title = element_text(size = 16)) +
  guides(fill = guide_colorbar(
    title.position = "top",      # Place the title above the color bar
      barwidth = 20,               # Adjust bar width for better horizontal layout
      barheight = 3,
    label.theme = element_text(angle = 45, hjust = 1)
  ))
  
p_obs_fire_rh_w
```

```{r}
p_obs_fire_t_rh |p_obs_fire_t_p | p_obs_fire_t_w | p_obs_fire_p_rh | p_obs_fire_p_w | p_obs_fire_rh_w

```

```{r}
p <- (p_r_strike_t_rh / p_r_strike_t_p / p_r_strike_t_w / p_r_strike_p_rh / p_r_strike_p_w / p_r_strike_rh_w)|(p_ignite_t_rh / p_ignite_t_p / p_ignite_t_w / p_ignite_p_rh / p_ignite_p_w / p_ignite_rh_w) | (p_fire_t_rh / p_fire_t_p / p_fire_t_w / p_fire_p_rh / p_fire_p_w / p_fire_rh_w)
#(p_obs_fire_t_rh / p_obs_fire_t_p / p_obs_fire_t_w / p_obs_fire_p_rh / p_obs_fire_p_w / p_obs_fire_rh_w)


```


```{r}
ggsave(file.path(figure_root, "Figure_8.png"), plot = p, width = 15, height = 30, dpi = 700, units = "in")
```