---
title: "Figure 4"
author: "Charlotte Uden"
date: "2026-04-20"
output: html_document
---

Reviewer comment:

The model(s) were fit and tested for 2002-2011 and 2012-2018. This makes sense from the standpoint of the sensor changes as described, however the results don’t mention any evaluation of RMSE, etc. outside of the fitting period so it’s hard to judge the efficacy of the results. Given the limited lightning data in Alaska, one solution could be to produce some evaluation statistics for the period not used in the fitting (i.e. model fit for 2002-2011 is evaluated also against the 2012-2018 period). Earlier lightning observations are also available starting in the early 1990s that could be helpful as another independent test set.

Our Response:
We agree that applying parameter estimates outside the fitting period would strengthen our assessment of model generalizability. We have taken your suggestion and, for the ‘best’ lightning and ignition models, applied parameters from each period to data from the alternate period to check for out-of-period performance. We calculated performance metrics from these new predictions and added them to Figures 2 and 3 for clear in-period and out-of-period comparison. Finally, we added sections 3.1.4 and 3.2.5, “Cross validation across time periods,” to the results section. 


Figure 4. Performance of ignition probability models across predictor groups, modeling approaches, and training periods. Panels show five evaluation metrics: (a) area under the receiver-operating characteristic curve (ROC AUC), (b) area under the precision-recall curve (PR AUC), (c) brier score, (d) calibration intercept, and (e) calibration slope. Points represent model estimates with horizontal bars indicating 95% bootstrap confidence intervals. Predictor groups are arranged on the y-axis, approach distinguished by point color (glmnet=green, Bayesian=orange) and training period distinguished by point shape (2002-2011=circle, 2012-2018=triangle). Each panel is divided into two sections: the upper facet (out-of-period test) shows performance of the climate + fuel moisture + vegetation Bayesian logistic regression when parameters estimated from one period are applied to data from the alternate period, while the lower facet (in-period test) shows performance evaluated on a held-out 20% test set from the same period used for training. In panels (a) and (b), higher values indicate stronger discrimination with dashed vertical lines marking no-skill thresholds (ROC AUC=0.5; PR AUC=1.55x10-4 for 2002-2011 and PR AUC=1.89x10-5 for 2012-2018). Panel (c) displays Brier scores on a reversed axis so that better performance appears on the right. Panels (d) and (e) assess calibration: a calibration intercept of 0 indicates unbiased predictions, and a slope of 1 indicates perfectly scaled probabilities (dashed vertical lines provide a reference).


```{r}
library(scales)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(scales)
library(stringr)
library(tidyverse)
library(cowplot)
library(ggpubr)
```


Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths 
```

```{r}
df <- read.csv(file.path(table_root, "ignition_efficiency_model_performance_time_transfer_test.csv"))
```

```{r}

# ── 1. Data preparation ───────────────────────────────────────────────────────

desired_order <- c(
  "climate", "fuel moisture", "vegetation",
  "climate & fuel moisture", "climate & vegetation",
  "fuel moisture & vegetation", "climate, fuel moisture & vegetation"
)

# In-period test set rows
df_inperiod <- df %>%
  filter(input == "test data (in-period)") %>%
  mutate(
    facet_group = "In-period",
    y_label     = as.character(predictors)
  )

# Cross-period transferability rows
df_transfer <- df %>%
  filter(input != "test data (in-period)") %>%
  mutate(
    facet_group = "Out-of-period",
    y_label     = as.character(y_label)  # already set as arrow labels in df_plot
  )

df_plot3 <- bind_rows(df_inperiod, df_transfer) %>%
  mutate(
    facet_group = factor(facet_group,
                         levels = c("Out-of-period", "In-period")),
    y_label = factor(y_label,
                     levels = c(
                       "2001-2012 \u2192 2012-2018",
                       "2012-2018 \u2192 2001-2012",
                       desired_order
                     )),
    approach       = factor(approach, levels = c("glmnet", "Bayesian")),
    training_years = factor(training_years, levels = c("2001-2012", "2012-2018"))
  )

# ── 2. Shared aesthetics ──────────────────────────────────────────────────────

color_vals3 <- c("glmnet" = "#d018d0", "Bayesian" = "#ffa908")
shape_vals3 <- c("2001-2012" = 16, "2012-2018" = 17)
dodge        <- position_dodge(width = 0.6)

base_theme3 <- theme_bw(base_size = 11) +
  theme(
    axis.title.y     = element_blank(),
    legend.position  = "none",
    strip.background = element_blank(),
    strip.text       = element_blank(),
    axis.text.x      = element_text(angle = 45, hjust = 1, vjust = 1),
    plot.margin      = ggplot2::margin(t = 2.5, r = 10, b = 2.5, l = 30)
  )

theme_with_strips3 <- base_theme3 +
  theme(
    strip.text.y     = element_text(size = 6, face = NULL,
                                    angle = 90,
                                    margin = margin(l = 4, r = 4)),
    strip.background = element_rect(fill = "white", color = "grey70"),
    strip.placement  = "outside"
  )

add_facet_scales3 <- function(p, show_strips = FALSE, right_axis = FALSE) {
  p2 <- p +
    facet_grid(facet_group ~ ., scales = "free_y", space = "free_y") +
    scale_color_manual(values = color_vals3, name = "Approach") +
    scale_shape_manual(values = shape_vals3,  name = "Training period")

  if (right_axis) {
    p2 <- p2 +
      scale_y_discrete(position = "right",
                       labels = function(x) str_wrap(x, width = 25)) +
      theme(plot.margin = ggplot2::margin(t = 10, r = 30, b = 10, l = 10))
  } else {
    p2 <- p2 +
      scale_y_discrete(labels = function(x) str_wrap(x, width = 25))
  }

  if (show_strips) {
    p2 + theme_with_strips3
  } else {
    p2 + base_theme3
  }
}

pts_eb3 <- function(xvar, xmin_var, xmax_var) {
  list(
    geom_point(aes(x = .data[[xvar]]),
               size = 2.8, alpha = 0.9, position = dodge),
    geom_errorbarh(aes(xmin = .data[[xmin_var]], xmax = .data[[xmax_var]]),
                   height = 0.25, alpha = 0.6, linewidth = 0.5,
                   position = dodge)
  )
}

# ── 3. Individual panels ──────────────────────────────────────────────────────

# (a) ROC AUC — strip labels ON, left y-axis
p_auc <- add_facet_scales3(show_strips = TRUE, right_axis = FALSE,
  p = ggplot(df_plot3, aes(y = y_label, color = approach, shape = training_years)) +
    pts_eb3("AUC", "AUC_CI_low", "AUC_CI_high") +
    geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey70") +
    scale_x_continuous(limits = c(0.3, 1.0),
                       breaks = seq(0.3, 1.0, by = 0.1),
                       labels = label_number(accuracy = 0.1)) +
    labs(x = "ROC AUC") +
    ggtitle("(a)")
)

# (b) PR AUC — no strips, RIGHT y-axis
p_pr_auc <- add_facet_scales3(show_strips = FALSE, right_axis = TRUE,
  p = ggplot(df_plot3, aes(y = y_label, color = approach, shape = training_years)) +
    pts_eb3("PR_AUC", "PR_AUC_CI_low", "PR_AUC_CI_high") +
    scale_x_continuous(limits = c(0.1, 1.0),
                       breaks = seq(0.1, 1.0, by = 0.1),
                       labels = label_number(accuracy = 0.1)) +
    labs(x = "PR AUC") +
    ggtitle("(b)")
)

# (c) Brier Score — reversed axis, no strips, left y-axis
p_brier <- add_facet_scales3(show_strips = FALSE, right_axis = FALSE,
  p = ggplot(df_plot3, aes(y = y_label, color = approach, shape = training_years)) +
    pts_eb3("Brier", "Brier_CI_low", "Brier_CI_high") +
    scale_x_reverse(breaks = seq(0.43, 0.13, by = -0.02),
                    labels = label_number(accuracy = 0.01),
                    limits = c(0.43, 0.13)) +
    labs(x = "Brier Score") +
    ggtitle("(c)")
)

# (d) Calibration Intercept — no strips, RIGHT y-axis
p_calib_int <- add_facet_scales3(show_strips = FALSE, right_axis = TRUE,
  p = ggplot(df_plot3, aes(y = y_label, color = approach, shape = training_years)) +
    pts_eb3("calib_intercept", "calib_intercept_CI_low", "calib_intercept_CI_high") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey70") +
    scale_x_continuous(breaks = seq(-2, 14, by = 2),
                       labels = label_number(accuracy = 1),
                       limits = c(-2, 14)) +
    labs(x = "Calibration Intercept") +
    ggtitle("(d)")
)

# (e) Calibration Slope — no strips, left y-axis
p_calib_slope <- add_facet_scales3(show_strips = FALSE, right_axis = FALSE,
  p = ggplot(df_plot3, aes(y = y_label, color = approach, shape = training_years)) +
    pts_eb3("calib_slope", "calib_slope_CI_low", "calib_slope_CI_high") +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey70") +
    scale_x_continuous(breaks = seq(-4, 15, by = 2),
                       labels = label_number(accuracy = 1),
                       limits = c(-4, 15)) +
    labs(x = "Calibration Slope") +
    ggtitle("(e)")
)

# ── 4. Shared horizontal legend ───────────────────────────────────────────────

legend_plot3 <- ggplot(df_plot3,
                       aes(x = AUC, y = y_label,
                           color = approach, shape = training_years)) +
  geom_point(size = 2.8) +
  scale_color_manual(values = color_vals3, name = "Approach") +
  scale_shape_manual(values = shape_vals3,  name = "Years") +
  guides(color = guide_legend(nrow = 1),
         shape = guide_legend(nrow = 1)) +
  theme_void() +
  theme(
    legend.position   = "bottom",
    legend.box        = "horizontal",
    legend.title      = element_text(size = 10),
    legend.text       = element_text(size = 9),
    legend.key.height = unit(0.5, "cm"),
    legend.key.width  = unit(1.0, "cm"),
    legend.spacing.x  = unit(2, "cm")
  )

shared_legend3 <- ggpubr::get_legend(legend_plot3)

# ── 5. Blank plot to fill grid ────────────────────────────────────────────────

blank_plot <- ggplot() + theme_void() + theme(plot.background = element_rect(fill = "white", color = NA))

# ── 6. Assemble final figure ──────────────────────────────────────────────────

fig_panels3 <- plot_grid(
  p_auc,    p_pr_auc,
  p_brier,  p_calib_int,
  p_calib_slope, blank_plot,
  ncol = 2,
  align = "h"
)

final_fig3 <- plot_grid(
  fig_panels3,
  shared_legend3,
  ncol = 1,
  rel_heights = c(1, 0.05)
)

final_fig3


ggsave(file.path(figure_root, "Figure_3.png"),
        plot = final_fig3, width = 10, height = 10, dpi = 1100, units = "in")

```




