---
title: "Figure_6"
author: "Charlotte Uden"
date: "2026-01-30"
output: html_document
---

Figure 6. Mapped predictions and observations of lightning strike rates, lightning ignition efficiency, and fire occurrence in Alaska, June-August 2002-2011. Predictions of lightning strike rates, ignition probability, and fire probability come from ERA5 climate and fuel moisture data, and NALCMS land cover data during the ten year period. Observed lightning strike rates are calculated from data collected by the Alaska Lightning Detection Network. Observed ignition rates are calculated as the number of strikes that caused a fire divided by the total number of strikes during the study period. Observed fire rates are calculated as the number of fire days divided by the total number of days. 

```{r}
library(tidyr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(terra)
library(lubridate)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
```

Alaska polygon
```{r}
# Get USA polygon
usa <- ne_states(country = "United States of America", returnclass = "sf")
# Filter to Alaska only
alaska <- usa[usa$name == "Alaska", ]
alaska <- st_transform(alaska, crs = 4326)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]
data_root <- config[["data_root"]]

# Use the values to build full paths 
```


#Read in calculations of observed lightning strike rate
```{r}
df_strikes <- read.csv(file.path(drivers_root, "lightning_prediction_drivers_2002_2011.csv"))[,2:13]

obs_strike <- df_strikes %>% dplyr::select(year,lon,lat,summer_monthly_strike_rate)

summary_obs_strike <- obs_strike %>%
  group_by(lat, lon) %>%
  summarise(
    obs_r_strike = mean(summer_monthly_strike_rate),
    .groups = "drop")
```


#Read in observations of fire occurance and predictions of r_strike, P_ignite, and P_fire
```{r}
df <- read.csv(file.path(drivers_root, "ignition_efficiency_drivers_2002_2011_10x.csv"))
```

#Calculate:
Mean predicted strike rate
Mean predicted ignition efficiency
Mean predicted fire probability
Mean observed fire rate (= #fire days / total # of days)
```{r}
summary_df <- df %>%
  group_by(lat, lon) %>%
  summarise(
    r_strike = mean(r_strike),
    P_ignite = mean(P_ignite_prob),
    fire_freq = mean(fire),
    n_fire = sum(fire), 
    P_fire = mean(i.P_fire),
    .groups = "drop")

#merge observed strike rate with summary df
summary_df <- merge(summary_df, summary_obs_strike, by=c("lat","lon"))

```

Get total number of strikes per gridcell for the study period (2001-2011)
monthly strike sums are calculated during preprocessing. See: /Lightning-Fire-Models-Uden-2026/Data_preprocessing/Lightning/3_lightning_era5climate_monthly_summaries.R
```{r}
#read in observed monthly strike count data
df_strikes_monthly <- read.csv(fil.path(data_root, "Alaska_era5_lightning_monthly_summaries_2001_2012.csv"))[,2:14]
df_strikes_monthly <- df_strikes_monthly %>% dplyr::select(year, month, lon, lat, strikes_monthly_sum)

#sum strikes
summary_obs_strike_sum <- df_strikes_monthly %>%
  group_by(lat, lon) %>%
  summarise(
    n_strikes = sum(strikes_monthly_sum))

#merge observed strike count with summary df
summary_df <- merge(summary_df, summary_obs_strike_sum, by=c("lat","lon"))

```
Calculate observed ignition event rate: the rate of lightning strikes that cause a fire (#fires/#strikes), per gridcell, across the entire study period (2001-2011)
```{r}
summary_df <- summary_df %>%
  mutate(obs_ignite = n_fire/n_strikes) %>%
  mutate(obs_ignite = replace_na(obs_ignite, 0))
```

Get min and max values for color plotting
```{r}
#global min and max fire probabilities/frequencies
min_value_fire <- min(c(summary_df$fire_freq, summary_df$P_fire), na.rm = TRUE)
max_value_fire <- max(c(summary_df$fire_freq, summary_df$P_fire), na.rm = TRUE)

#global min and max strike rates
min_value_strike <- min(c(summary_df$obs_r_strike, summary_df$r_strike), na.rm = TRUE)
max_value_strike <- max(c(summary_df$obs_r_strike, summary_df$r_strike), na.rm = TRUE)

#global min and max ignition probabilities/frequencies
min_value_ignite <- min(c(summary_df$obs_ignite, summary_df$P_ignite), na.rm = TRUE)
max_value_ignite <- max(c(summary_df$obs_ignite, summary_df$P_ignite), na.rm = TRUE)
```

Make maps!

```{r}
p_strike <- ggplot(summary_df, aes(x = lon, y = lat, fill = r_strike)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_strike, max_value_strike), 
    name = expression(atop("Predicted lightning strike rate", "strikes/month/km"^2)),
    guide = guide_colorbar(direction = "horizontal", title.position = "top"))  +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(a)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_strike
```


```{r}
p_strike_obs <- ggplot(summary_df, aes(x = lon, y = lat, fill = obs_r_strike)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_strike, max_value_strike), 
    name = expression(atop("Observed lightning strike rate", "strikes/month/km"^2)),
    guide = guide_colorbar(direction = "horizontal", title.position = "top"))  +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(d)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_strike_obs
```

```{r}
p_ignite <- ggplot(summary_df, aes(x = lon, y = lat, fill = P_ignite)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    #limits = c(min_value_ignite, max_value_ignite), 
    name = expression(Predicted~ignition~efficiency),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(b)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_ignite
```

```{r}
p_ignite_obs <- ggplot(summary_df, aes(x = lon, y = lat, fill = obs_ignite)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    #limits = c(min_value_ignite, max_value_ignite), 
    #name = expression(Observed~ignition~rate),
    name = expression(atop("Observed ignition rate", "(fire count / strike count)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(e)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_ignite_obs
```


```{r}
p_fire <- ggplot(summary_df, aes(x = lon, y = lat, fill = P_fire)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_fire, max_value_fire), 
    name = expression(Predicted~fire~probability),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_fire
```

```{r}
p_obs_fire <- ggplot(summary_df, aes(x = lon, y = lat, fill = fire_freq)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_fire, max_value_fire), 
    #name = expression(Observed~fire~rate),
    name = expression(atop("Observed fire rate", "(fire days / total days)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(f)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size = 8),
    axis.title.y = element_text(size = 8),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(size = 8),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = "white", color = "white", size = 0.2),
    legend.position = c(0.6, 0.15))

p_obs_fire
```

```{r}
blank_plot <- ggplot() + theme_void()

p <- grid.arrange(p_strike, p_ignite, p_fire, ncol=1)

p <- grid.arrange(p_strike, p_ignite, p_fire, p_obs_fire, ncol=1)

p <- grid.arrange(p_strike, p_ignite, p_fire, p_strike_obs, p_ignite_obs, p_obs_fire, ncol=3)
```

```{r}
ggsave(file.path(figure_root, "Figure_6.png"), plot = p, width = 15, height = 10, dpi = 1100, units = "in")
```
