---
title: "Figure_7"
author: "Charlotte Uden"
date: "2026-01-30"
output: html_document
---

Figure 7. Mapped predictions and observations of lightning strike rates, lightning ignition efficiency, and fire occurrence in Alaska, June-August 2002-2011. The top row contains model predictions of (a) lightning strike rates, (b) ignition probability, and (c) fire probability, estimated from ERA5 climate and fuel moisture data, and NALCMS land cover data and are summarized as the mean prediction per gridcell across the ten year period. The middle row contains summaries of observed data, including (d) mean lightning strike rate, (e) ignition rates (number of strikes that caused a fire / total number of strikes), and (f) fire rate (number of fire days / total number of days). Lightning and wildfire data come from the Alaska Lightning Detection Network and the Alaska-Yukon-Northwest Territories Fire Emissions Database. The bottom row contains per-gridcell calculations of mean bias for (g) lightning strike rate (seasonal predicted - observed strikes/km2/month) and (h) fire probability (daily predicted probability - observed 0/1 outcome). Purple gridcells indicate underprediction and orange gridcells indicate overprediction. Topographic lines downloaded using the geodata R package [56] are also displayed. 


```{r}
library(tidyr)
library(patchwork)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(terra)
library(lubridate)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

#config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]

# Use the values to build full paths 
```

Alaska polygon
```{r}
# Get USA polygon
usa <- ne_states(country = "United States of America", returnclass = "sf")
# Filter to Alaska only
alaska <- usa[usa$name == "Alaska", ]
alaska <- st_transform(alaska, crs = 4326)
```

#Read in calculations of observed lightning strike rate
```{r}
df_strikes <- read.csv(file.path(drivers_root, "lightning_prediction_drivers_2002_2011.csv"))[,2:13]

obs_strike <- df_strikes %>% dplyr::select(year,lon,lat,summer_monthly_strike_rate)

summary_obs_strike <- obs_strike %>%
  group_by(lat, lon) %>%
  summarise(
    obs_r_strike = mean(summer_monthly_strike_rate),
    .groups = "drop")
```


#Read in observations of fire occurance and predictions of r_strike, P_ignite, and P_fire
```{r}
df <- read.csv(file.path(drivers_root, "ignition_efficiency_drivers_2002_2011_10x.csv"))
```

#Calculate:
Mean predicted strike rate
Mean predicted ignition efficiency
Mean predicted fire probability
Mean observed fire rate (= #fire days / total # of days)
```{r}
summary_df <- df %>%
  group_by(lat, lon) %>%
  summarise(
    r_strike = mean(r_strike),
    P_ignite = mean(P_ignite_prob),
    fire_freq = mean(fire),
    n_fire = sum(fire), 
    P_fire = mean(i.P_fire),
    .groups = "drop")

#merge observed strike rate with summary df
summary_df <- merge(summary_df, summary_obs_strike, by=c("lat","lon"))

```

#Calculate:
error for strike rate prediciton
error for fire prediciton
```{r}
# Aggregate df to monthly level (one row per gridcell per month/year)
df_monthly <- df %>%
  group_by(lat, lon, year, month) %>%
  summarise(
    r_strike_pred = first(r_strike),  # already a monthly mean, just take first value
    .groups = "drop"
  )

# Merge with observed monthly strike rate
df_monthly <- df_monthly %>%
  left_join(
    df_strikes %>% dplyr::select(year, lon, lat, summer_monthly_strike_rate),
    by = c("year", "lon", "lat")
  )

gridcell_bias_strike <- df_monthly %>%
  group_by(lat, lon) %>%
  #filter(sum(!is.na(summer_monthly_strike_rate)) >= 3) %>%
  summarise(
    mean_strike_bias = mean(r_strike_pred - summer_monthly_strike_rate, na.rm = TRUE),
    .groups = "drop"
  )

#calculate bias for fire prediction
gridcell_bias_fire <- df %>%
  group_by(lat, lon) %>%
  summarise(
    bias_fire = mean(i.P_fire - fire, na.rm = TRUE),
    .groups = "drop"
  )

summary_df <- summary_df %>%
  left_join(gridcell_bias_fire, by = c("lat", "lon"))%>%
  left_join(gridcell_bias_strike, by = c("lat", "lon"))
```

Get total number of strikes per gridcell for the study period (2001-2011)
```{r}
#read in observed monthly strike count data
df_strikes_monthly <- read.csv(fil.path(data_root, "Alaska_era5_lightning_monthly_summaries_2001_2012.csv"))[,2:14]
df_strikes_monthly <- df_strikes_monthly %>% dplyr::select(year, month, lon, lat, strikes_monthly_sum)

#sum strikes
summary_obs_strike_sum <- df_strikes_monthly %>%
  group_by(lat, lon) %>%
  summarise(
    n_strikes = sum(strikes_monthly_sum))

#merge observed strike count with summary df
summary_df <- merge(summary_df, summary_obs_strike_sum, by=c("lat","lon"))

```

Calculate observed ignition rate: the rate of lightning strikes that cause a fire (#fires/#strikes), per gridcell, across the entire study period (2001-2011)
```{r}
summary_df <- summary_df %>%
  mutate(obs_ignite = n_fire/n_strikes) %>%
  mutate(obs_ignite = replace_na(obs_ignite, 0))
```

Get min and max values for color plotting
```{r}
#global min and max fire probabilities/frequencies
min_value_fire <- min(c(summary_df$fire_freq, summary_df$P_fire), na.rm = TRUE)
max_value_fire <- max(c(summary_df$fire_freq, summary_df$P_fire), na.rm = TRUE)

#global min and max strike rates
min_value_strike <- min(c(summary_df$obs_r_strike, summary_df$r_strike), na.rm = TRUE)
max_value_strike <- max(c(summary_df$obs_r_strike, summary_df$r_strike), na.rm = TRUE)

#global min and max ignition probabilities/frequencies
min_value_ignite <- min(c(summary_df$obs_ignite, summary_df$P_ignite), na.rm = TRUE)
max_value_ignite <- max(c(summary_df$obs_ignite, summary_df$P_ignite), na.rm = TRUE)
```

Make maps!

```{r}
p_strike <- ggplot(summary_df, aes(x = lon, y = lat, fill = r_strike)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_strike, max_value_strike), 
    name = expression(atop("Predicted lightning", "strikes/month/km"^2)),
    guide = guide_colorbar(direction = "horizontal", title.position = "top"))  +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(a)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.16))

p_strike
```


```{r}
p_strike_obs <- ggplot(summary_df, aes(x = lon, y = lat, fill = obs_r_strike)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_strike, max_value_strike), 
    name = expression(atop("Observed lightning", "strikes/month/km"^2)),
    guide = guide_colorbar(direction = "horizontal", title.position = "top"))  +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(d)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.16))

p_strike_obs
```



```{r}
max_abs_bias <- max(abs(summary_df$mean_strike_bias), na.rm = TRUE)

p_strike_bias <- ggplot(summary_df, aes(x = lon, y = lat, fill = mean_strike_bias)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  # scale_fill_viridis_c(
  #   option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
  #   direction = -1,
  #   name = expression(atop("Mean Bias", "strikes/month/km"^2)),
  #   guide = guide_colorbar(direction = "horizontal", title.position = "top"))  +
   scale_fill_distiller(
  palette = "PuOr",
  limits = c(-max_abs_bias, max_abs_bias),
  name = expression(atop("Mean Bias", "strikes/month/km"^2)),
  guide = guide_colorbar(direction = "horizontal", title.position = "top")
) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(g)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.16))

p_strike_bias
```

```{r}
p_ignite <- ggplot(summary_df, aes(x = lon, y = lat, fill = P_ignite)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    #limits = c(min_value_ignite, max_value_ignite), 
    #name = expression(Predicted~ignition~efficiency),
    name = expression(atop("Predicted ignition efficiency", "P(ignition | lightning)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(b)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.margin = margin(t = 5, r = 15, b = 5, l = 10), 
    legend.position = c(0.65, 0.16))

p_ignite
```

```{r}
p_ignite_obs <- ggplot(summary_df, aes(x = lon, y = lat, fill = obs_ignite)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    #limits = c(min_value_ignite, max_value_ignite), 
    #name = expression(Observed~ignition~rate),
    name = expression(atop("Observed ignition rate", "(fire count / strike count)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(e)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.margin = margin(t = 5, r = 20, b = 5, l = 10), 
    legend.position = c(0.65, 0.15))

p_ignite_obs
```


```{r}
p_fire <- ggplot(summary_df, aes(x = lon, y = lat, fill = P_fire)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_fire, max_value_fire), 
    #name = expression(Predicted~fire~probability),
    name = expression(atop("Predicted fire probability", "P(fire | lightning, ignition efficiency)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.margin = margin(t = 5, r = 20, b = 5, l = 10), 
    legend.position = c(0.65, 0.16))

p_fire
```

```{r}
p_obs_fire <- ggplot(summary_df, aes(x = lon, y = lat, fill = fire_freq)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_viridis_c(
    option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
    direction = -1,
    limits = c(min_value_fire, max_value_fire), 
    #name = expression(Observed~fire~rate),
    name = expression(atop("Observed fire rate", "(fire days / total days)")),
    guide = guide_colorbar(direction = "horizontal", title.position = "top")
  ) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(f)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.margin = margin(t = 5, r = 20, b = 5, l = 10), 
    legend.position = c(0.65, 0.16))

p_obs_fire
```


```{r}
max_abs_bias <- max(abs(summary_df$bias_fire), na.rm = TRUE)

p_bias_fire <- ggplot(summary_df, aes(x = lon, y = lat, fill = bias_fire)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  # scale_fill_viridis_c(
  #   option = "plasma",  # Default Viridis color scale (use "cividis" or "plasma" if needed)
  #   #direction = -1,
  #   name = expression(ROC~AUC),
  #   guide = guide_colorbar(direction = "horizontal", title.position = "top")
  # ) +
  scale_fill_distiller(
  palette = "PuOr",
  limits = c(-max_abs_bias, max_abs_bias),  # symmetric around 0
  name = expression(atop("Mean Bias", "(fire probability)")),
  #name = "Mean bias\n([predicted - observed)",
  guide = guide_colorbar(direction = "horizontal", title.position = "top")
) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(i)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.17))

p_bias_fire
```

Elevation plot
```{r}
library(geodata)
library(terra)

# Download SRTM elevation data for Alaska
# Alaska spans multiple tiles so we download by country

# 1. Download and crop
elev <- geodata::elevation_global(res = 10, path = tempdir())
alaska_ext <- ext(-170, -130, 51, 72)
elev_alaska <- crop(elev, alaska_ext)

# 2. Mask to Alaska polygon
alaska_vect <- terra::vect(alaska)
elev_masked <- terra::mask(elev_alaska, alaska_vect)

# 3. Convert to data frame — use elev_masked instead of elev_alaska
elev_df <- as.data.frame(elev_masked, xy = TRUE)
names(elev_df) <- c("lon", "lat", "elevation")
elev_df <- elev_df %>% filter(!is.na(elevation), elevation >= 0)

p_elevation <- ggplot() +
  geom_contour(data = elev_df,
               aes(x = lon, y = lat, z = elevation),
               color = "black",
               linewidth = 0.1,
               breaks = seq(0, 3000, by = 250)) +  # contour interval in metres
  geom_sf(data = alaska, fill = NA, color = "black",
          linewidth = 0.3, inherit.aes = FALSE) +
  xlim(-170, -130) +
  ylim(51, 72) +
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(h)") +
  theme_minimal() +
  theme(
    axis.title.x     = element_text(size=11),
    axis.title.y     = element_text(size=11),
    axis.text.y      = element_text(size=11),
    axis.text.x      = element_text(size=11),
    panel.grid.major = element_line(color = "darkgray", size = 0.2),
    panel.grid.minor = element_line(color = "darkgray", size = 0.2)
  )

p_elevation
```

```{r}
max_abs_bias <- max(abs(summary_df$bias_fire), na.rm = TRUE)

p_bias_fire_elev <- ggplot(summary_df, aes(x = lon, y = lat, fill = bias_fire)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
  scale_fill_distiller(
  palette = "PuOr",
  limits = c(-max_abs_bias, max_abs_bias),  # symmetric around 0
  name = expression(atop("Mean Bias", "(fire probability)")),
  guide = guide_colorbar(direction = "horizontal", title.position = "top")
) +
  geom_contour(data = elev_df,
               aes(x = lon, y = lat, z = elevation),
               color = "black",
               linewidth = 0.1,
               breaks = seq(0, 3000, by = 250),
               inherit.aes = FALSE) +  
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(h)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.17))

p_bias_fire_elev

```

```{r}
max_abs_bias <- max(abs(summary_df$mean_strike_bias), na.rm = TRUE)

p_strike_bias_elev <- ggplot(summary_df, aes(x = lon, y = lat, fill = mean_strike_bias)) +
  geom_raster() +
  xlim(-170, -130) +
  geom_sf(data = alaska, fill = NA, color = "black", linewidth = 0.3, inherit.aes = FALSE) +
   scale_fill_distiller(
  palette = "PuOr",
  limits = c(-max_abs_bias, max_abs_bias),
  name = expression(atop("Mean Bias", "strikes/month/km"^2)),
  guide = guide_colorbar(direction = "horizontal", title.position = "top")
) +
  geom_contour(data = elev_df,
               aes(x = lon, y = lat, z = elevation),
               color = "black",
               linewidth = 0.1,
               breaks = seq(0, 3000, by = 250),
               inherit.aes = FALSE) +  
  xlab("Lon") +
  ylab("Lat") +
  ggtitle("(g)") +
  theme_minimal() +
  theme(legend.position="bottom") +
  theme(
    axis.title.x = element_text(size=11),
    axis.title.y = element_text(size=11),
    axis.text.y = element_text(size=11),
    axis.text.x = element_text(size=11),
    legend.text = element_text(size = 8, angle = 45, hjust = 1),
    legend.title = element_text(size = 8, hjust = 0),
    panel.grid.major = element_line(color = "darkgray", size = 0.2), 
    panel.grid.minor = element_line(color = "darkgray", size = 0.2),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.position = c(0.65, 0.16))

p_strike_bias_elev
```

```{r}
#blank_plot <- ggplot() + theme_void()

tight_margin <- theme(plot.margin = margin(t = 2, r = 2, b = 2, l = 2))

p_strike      <- p_strike      + tight_margin
p_ignite      <- p_ignite      + tight_margin
p_fire        <- p_fire        + tight_margin
p_strike_obs  <- p_strike_obs  + tight_margin
p_ignite_obs  <- p_ignite_obs  + tight_margin
p_obs_fire    <- p_obs_fire    + tight_margin

library(patchwork)

top_rows <- (p_strike | p_ignite | p_fire) /
            (p_strike_obs | p_ignite_obs | p_obs_fire)

bottom_row <- (p_strike_bias_elev | p_bias_fire_elev)

final <- top_rows / bottom_row +
  plot_layout(heights = c(1, 1, 1.5))
```

```{r}
ggsave(file.path(figure_root, "Figure_6.png"), plot = final, width = 6, height = 12, dpi = 1100, units = "in")
```
