---
title: "Transforming Observations"
author: "Charlotte Uden"
date: "2026-01-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(terra)
library(sf)
library(lubridate)
library(MASS)
library(fields)
library(lubridate)
library(geosphere)
library(ggspatial)
library(viridis)
library(rlang)  
```

Load configuration files
- location of model drivers, rds files (containing models), tables of parameter estimates and performance metrics, and figures. 
```{r}
read_properties <- function(file_path) {
  lines <- readLines(file_path)
  lines <- lines[grepl("=", lines)]  # only keep lines with key-value pairs
  key_vals <- strsplit(lines, "=")
  props <- setNames(
    trimws(sapply(key_vals, `[`, 2)),
    trimws(sapply(key_vals, `[`, 1))
  )
  return(props)
}

config <- read_properties("/path/to/config/file/my_config.properties")  # path to your config file

# Extract values from config
drivers_root <- config[["drivers_root"]]
rds_root <- config[["rds_root"]]
table_root <- config[["table_root"]]
predictions_root <- config[["predictions_root"]]
figure_root <- config[["figure_root"]]
data_root <- config[["data_root"]]
```

This script converts the location and time of observed lightning-caused wildfire in Alaska (2001-2018) to the ERA5 grid. 
This is for comparison with predictions of lightning caused fires, which are defined as:

P_fire,i,t = P(>= 1 lightning-caused fire in grid cell i on day t)

The observations need to be converted to a binary event: fire occurred or did no occur for comparison with a bernoulli probability (not a rate or count). This is defined on a grid cell x day. So, the observation target should be defined like so:

"Did at least one lightning-caused fire occur in grid cell i on day t?"

So, we must convert the observations to match this by 
1. snap fire points to model grid (ERA5 grid)
2. collapse to binary grid cell x day. 

Alaska polygon
```{r}
library(rnaturalearth)
library(rnaturalearthdata)

# Get USA polygon
usa <- ne_states(country = "United States of America", returnclass = "sf")
# Filter to Alaska only
alaska <- usa[usa$name == "Alaska", ]
alaska <- st_transform(alaska, crs = 4326)

```

Read in lightning-caused fire observations

```{r}
#raw fire location data:
original_sf <- st_read(file.path(data_root, "Alaska_ignition_locations_20012018/ignitions_20012018_AK.shp"))
original_crs <- st_crs(original_sf)

#fires matched with a lightning strike:
fire1 <- read.csv(file.path(data_root, "binary_fire_lightning_matches_2002_2011.csv"))
fire2 <- read.csv(file.path(data_root, "binary_fire_lightning_matches_2012_2018.csv"))

fire <- rbind(fire1, fire2)
colnames(fire) <- c("X","lightning_date","lat","lon","caused_fire")

fire$lightning_date <- as.Date(fire$lightning_date, format = "%Y-%m-%d")

fire_sf <- st_as_sf(
  fire %>% filter(caused_fire == 1),   
  coords = c("lon", "lat"),
  crs = 4326, #use crs from shapefile we got the original data from (above)
  remove = FALSE
)

#to map fire locations, we need the crs to match the alaska ploygon
fire_sf <- st_transform(fire_sf, st_crs(alaska))

ggplot() +
  geom_sf(data = alaska, fill = "lightgray") +
  geom_sf(data = fire_sf, color = "red", size = 2, alpha=0.1) +
  theme_minimal() +
  labs(title = "Fire start locations",
       x = "Longitude", y = "Latitude")
```



ERA5 grid
```{r}
grid <- read.csv(file.path(data_root, "era5_grid.csv"))

grid_sf <- st_as_sf(
  grid %>% mutate(grid_id = X.1),   # or row_number() if you prefer
  coords = c("X", "Y"),
  crs = 4326,
  remove = FALSE
)

ggplot() +
  geom_sf(data = alaska, fill = "lightgray") +
  geom_sf(data = grid_sf, size=2) +
  coord_sf(xlim = c(-180, -120), ylim = c(47, 80)) +
  theme_minimal() +
  labs(title = "ERA5 grid",
       x = "Longitude", y = "Latitude")

# Transform grid to match fires CRS
grid_sf_firecrs <- st_transform(grid_sf, st_crs(fire_sf))
```

1. Snap fire points to the model (ERA5) grid. For each lightning-caused fire, assign it to a grid cell and extract the ignition date (day). 
```{r}
# Find nearest ERA5 grid point for each fire
nearest_idx <- st_nearest_feature(fire_sf, grid_sf_firecrs)

fires_gridded <- fire_sf %>%
  mutate(grid_id = grid_sf_firecrs$grid_id[nearest_idx]) %>%
  st_drop_geometry()

```

2. Aggregate to binary grid cell x day for occurrance
```{r}
obs_grid_day <- fires_gridded %>%
  group_by(grid_id, lightning_date) %>%
  summarise(
    fire = 1L,          # by construction, at least one fire occurred
    n_fires = n(),      # optional: keep counts for diagnostics
    .groups = "drop"
  )

```

3. Create a complete grid / day table that incldues zeros. 
```{r}
date_min <- min(fires_gridded$lightning_date, na.rm = TRUE)
date_max <- max(fires_gridded$lightning_date, na.rm = TRUE)

all_days <- seq.Date(date_min, date_max, by = "day")

obs_grid_day_full <- tidyr::expand_grid(
  grid_id = grid_sf$grid_id,
  lightning_date = all_days
) %>%
  left_join(obs_grid_day %>% dplyr::select(grid_id, lightning_date, fire),
            by = c("grid_id", "lightning_date")) %>%
  mutate(fire = ifelse(is.na(fire), 0L, fire))

head(obs_grid_day_full)
```

4. Sanity checks

```{r}
# How many fires got a grid cell?
sum(is.na(fires_gridded$grid_id))

# How many unique gridcells have >=1 fire?
n_distinct(obs_grid_day$grid_id)

# Any days with multiple fires in same grid cell?
obs_grid_day %>% filter(n_fires > 1) %>% arrange(desc(n_fires)) %>% head()

```
5. Merge with grid lat lon by grid id
```{r}
obs_grid_day_full_sf <- obs_grid_day_full %>%
  left_join(
    grid_sf_firecrs %>% dplyr::select(grid_id, geometry),
    by = "grid_id"
  ) %>%
  sf::st_as_sf()

#get explicit lat lon values
obs_grid_day_full_sf <- obs_grid_day_full_sf %>%
  mutate(
    lon = sf::st_coordinates(.)[, 1],
    lat = sf::st_coordinates(.)[, 2]
  )

str(obs_grid_day_full_sf)

#plot just 2002
obs_2002 <- obs_grid_day_full_sf %>%
  filter(lubridate::year(lightning_date) == 2002)

ggplot() +
  geom_sf(data = alaska, fill = "lightgray") +
  geom_sf(
    data = obs_2002,
    aes(color = factor(fire), alpha=0.1, size=2)
  ) +
  coord_sf(xlim = c(-180, -120), ylim = c(47, 80)) +
  theme_minimal() +
  labs(
    title = "Binary Lightning-Caused Fire Observations (2001)",
    color = "Fire"
  )

#plot only fires for 2002
obs_2002_fire <- obs_grid_day_full_sf %>%
  filter(lubridate::year(lightning_date) == 2002, fire == 1)

ggplot() +
  geom_sf(data = alaska, fill = "lightgray") +
  geom_sf(
    data = obs_2002_fire,
    color = "red",
    size = 2
  ) +
  coord_sf(xlim = c(-180, -120), ylim = c(47, 80)) +
  theme_minimal() +
  labs(
    title = "Lightning-Caused Fire Grid Cells (2001)"
  )
```
Final product:

obs_grid_day_full_sf contains one row per grid cell per day, with:

grid_id: unique ERA5 grid cell identifier

fire_date: calendar day (summer season)

fire: binary indicator: 1 if >= lightning-caused fire occurred in that frid cell on that day, 0 no fire

geometry: point geometry of the ERA5 grid cell (for mapping / spatial joins)

My model says: "Given the climate and lightning regime, what is the probability that at least one lightning-caused fire occurs here today?"

Save dataset:
```{r}
st_write(
  obs_grid_day_full_sf,
  file.path(model_drivers, "observed_binary_fire_data_2002-2018.shp"),
  delete_layer = TRUE
)
```
